### 2019-05-01 Paper_reorganized.Rmd

---
title             : "Quantitative involvement moderates associations between qualitative parenting behaviors and children's emotional adjustment development"
shorttitle        : "INVOLVEMENT AND EMOTIONAL ADJUSTMENT"

author: 
  - name          : "Caspar J. van Lissa"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email         : "c.j.vanlissa@uu.nl"
  - name          : "Renske Keizer"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Utrecht University, Methodology & Statistics"
  - id            : "2"
    institution   : "Erasmus University Rotterdam, Erasmus School of Social and Behavioral Sciences, department of Public Administration and Sociology"

author_note: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["DSEM_2017.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : yes

class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library("papaja")
library(motley)
source("rfunctions.R")
load("results.RData")
load("table_coefficients.RData")
mask <- TRUE
masked <- function(text){
  ifelse(mask, "[masked]", text)
}
```

```{r analysis_preferences}
# Seed for random number generation
set.seed(42)
```

Emotional adjustment is defined as the absence of internalizing symptoms and manifest signs of emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001]. This important developmental outcome is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013; @aldaoEmotionregulationStrategiesPsychopathology2010]. According to the fatherhood literature, emotional adjustment development is one domain in which fathers might play a unique role, because fathers encourage children to challenge themselves, in play and in life [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000]. The classical notion of the "attachment relationship" is rooted in a balance between security and exploration [@grossmannDeveloppementAttachementAdaptation1998]. Fatherhood researchers have posited that, whereas the mother-child attachment relationship primarily provides warmth and security, exploration might be somewhat more central to the father-child relationship, which can has thus been construed as an "activation relationship" [@paquetteFatherChildActivationRelationship2013]. By encouraging children to venture outside their comfort zone, fathers are thought to stimulate a broad spectrum of emotions, which expands children's emotional repertoire, and engages their developing regulatory abilities.

## Fatherhood theory

The past decades have seen major strides in the consolidation of theories to guide fatherhood research. Initially, such research was inspired by two societal shifts: Increasing divorce rates, and mothers' greater labor force participation. Both shifts impacted the extent of fathers' involvement in childrearing. Early research focused primarily on the *quantity* of paternal involvement: Whether fathers were present or absent in divorced families, which constitutes a dichotomous operationalization of quantitative involvement, and how much they contributed to the work of childcare in intact families, i.e., a continuous operationalization [@lambBiosocialPerspectivePaternal1987]. Over time, the notion of the father as a coparent gained ground [@pleckFatherhoodIdealsUnited1997], according to which fathers share the responsibilities of childrearing with mothers, and might, to some extent, adopt roles that were historically fulfilled by mothers. To keep pace with changing conceptualizations of fatherhood, research on father involvement evolved to encompass *qualitative* dimensions as well, including warmth and control [or: demandingness; @maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].

Although it is now widely acknowledged that both the quality and quantity of parental involvement are relevant factors in child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated. This might be, in part, because research interests have historically varied across discipline lines: Developmental psychology has focused primarily on how qualitative parenting behaviors influence child development, whereas sociology and gender studies have focused more on involvement quantity, and its implications for maternal well-being and gender equality [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014]. The present study attempts to bridge these approaches by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.

## Qualitative parenting and emotional adjustment

### Parenting effects

The standard family environment model posits that the quality of parents' involvement is a proximal mechanism, shaping children's development [@amatoParentalDivorceMarital2008]. Qualitative parenting is typically examined in terms of the constituent dimensions of the parenting styles taxonomy [@baumrindChildCarePractices1967]: Responsiveness and demandingness [@maccobySocializationContextFamily1983]. Responsiveness refers to warm and supportive parenting, which substantial evidence indicates is positively associated with children's socio-emotional adjustment [@khalequePerceivedParentalAcceptanceRejection2002; @morrisRoleFamilyContext2007].

The literature on links between demandingness and emotional adjustment is more limited, and mixed, than the literature on responsiveness [@morrisRoleFamilyContext2007]. This appears to be, in part, because some forms of demandingness help children meet societal expectations, whereas other forms are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013]. Demandingness encompasses several parenting practices related to setting and enforcing expectations for children's behavior. Two specific aspects of demandingness, investigated in the present study, are consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments. Consistency has been found to correlate strongly with emotional adjustment [@brandPerceivedParentalInconsistency1990]. Inductive reasoning is thought to promote the internalization of parental expectations, and is integral to children's development of moral emotions [@hoffmanEmpathyMoralDevelopment2000]. It will thus likely play a positive role in general emotional development as well.

Negative parenting is an additional qualitative parenting dimension which is regularly included in the parenting styles literature, and appears to be negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999]. We operationalized negative parenting as one of its specific subcomponents: parental anger. It is thought that the anger parents express in interactions with children reflects a faillure to regulate their own negative emotions, models children's use of anger in communication, and undermines children's emotion regulation [@scaramellaParentingChildrenDistress2008].

Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' influence [@flouriFathersBehaviorsChildren2010]. Although some research suggests that both parents' qualitative parenting behaviors influence children in similar ways [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique roles of mothers versus fathers. Whether for evolutionary or cultural reasons [@bogelsDoesFatherKnow2011], mothers and fathers sometimes appear to fulfill somewhat specialized roles in terms of qualitative parenting: The aforementioned notions of "attachment" and "activation" relationships posit that mothers empathize more with their children, and  provide more warmth and support, whereas fathers activate children by encouraging them to take risks and overcome obstacles, and live up to certain standards for behavior [@majdandzicFathersChallengingParenting2014; @paquetteTheorizingFatherchildRelationship2004]. Empirical research supports these unique roles for mothers and fathers in children's emotional adjustment development: Mothers tend to encourage the expression of emotions, and respond warmly and supportively, whereas fathers structure and sanction emotion expressions to a greater extent [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999]. Moreover, mothers' supportive behavior was found to be a more salient socialization mechanism than control for emotional adjustment in adolescence, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a]. This differentiated pattern was also found with regard to younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003]. Furthermore, there is some evidence that mothers' negative parenting has a stronger negative impact upon children's emotional adjustment than fathers' negative parenting [@changHarshParentingRelation2003]; perhaps because this behavior is incongruous with mothers' prototypical role as an attachment figure. Based on these findings, we might expect emotional adjustment to be more strongly associated with maternal warmth, and with paternal demanding behaviors, such as consistency and inductive reasoning.

### Child effects

According to the transactional view of family relationships, parents and children reciprocally influence one another [@kerrPerceivedParentingStyle2012; @amatoParentalDivorceMarital2008]. This view of the family as a dynamic system whose members reciprocally influence one another has become widely accepted [@kerrPerceivedParentingStyle2012]. Nevertheless, examining bi-directional effects is not yet standard practice in the childhood literature [@flouriFathersBehaviorsChildren2010]. Available empirical evidence suggests that children's adjustment problems elicit less warmth, more demandingness, and more negative parenting from parents [@gadeyneLongitudinalRelationsParenting2004; @nepplIntergenerationalContinuityParenting2009; @lenguaBidirectionalAssociationsTemperament2005]. In the adolescent literature, bi-directional effects are more commonly examined. This might be because the notion of children influencing parents is more salient in adolescence, when parent-child relationships become increasingly egalitarian. Findings from the adolescent literature are broadly in line with findings in childhood; suggesting that emotional adjustment problems predicts decreased parental support, and greater control [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014; @wijsbroekDirectionEffectsPerceived2011]. Thus, parents' responses to maladjustment appear to be a combination of punitive responses, including decreased support and increased negative parenting, and corrective responses, such as increased demandingness. Moreover, parents' responses to adjustment problems are contingent on children's age: Because younger children are perceived to be less competent than older children, mothers respond to their misbehavior more inductively, and less punitively [@dixAffectiveOrganizationParenting1991]. We might thus expect emotional adjustment to positively predict parental warmth, and negatively predict consistency, inductive reasoning, and anger.

### Assessing the direction of effects

Methods commonly used to examine direction of effects between parents and children have been criticized for yielding biased results if there are stable between-family differences in the variables studied [@hamakerCritiqueCrosslaggedPanel2015]. Recent methodological innovations provide better estimates of dynamic reciprocal processes within families, by controlling for stable between-family differences: The random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018].

Several studies have re-examined the direction of effects between parents and children using these innovative methods, but mostly in adolescent samples. One such study found that predictive effects between parenting and delinquency adjustment that were apparent using classical methods, were found to be explained by correlations between stable between-family differences in parenting and adjustment, once these components were included in the model [@keijsersParentalMonitoringAdolescent2016]. Another study examined the interplay between maternal and paternal qualitative parenting dimensions and adolescents' emotional adjustment, and similarly found that most associations manifested as between-family level correlations. Furthermore, child effects were found to substantially outnumber parenting effects [@vanlissaRoleFathersMothers2018a]. The authors suggested that this might be due to the relative equality of adolescent-parent relationships, as compared to parent-child relationships. If this interpretation is correct, we might expect to find more prominent parenting effects in the present study, with a younger sample. These preliminary findings highlight the fact that the direction of effects between parents and children, while controlling for stable between-family differences, remains an important issue to be examined. At this stage, however, it is difficult to formulate specific hypotheses about directionality, and the level at which associations between parenting and child development should occur.

<!-- Kroll: No evidence for effect of quantitative involvement This might help explain why some studies, focusing on absolute involvement, have concluded that fathers' quantitative involvement does not predict children's adjustment when controlling for mothers' involvement>> Kroll actually used absolute involvement and "who does most"-questions, which measure relative involvement -->

<!--
Better quality parental relationships are usually associated with more involved fathering (Belsky 1984;
Bouchard & Lee 2000; Coiro & Emery 1998; McBride & Mills 1993; Verhoeven et al. 2007). Further, when
relationships are better, this is also evident in more positive or sensitive parenting styles and more cooperative
and supportive co‑parenting (Floyd, Gilliom & Costigan 1998; Verhoeven et al. 2007).
-->

## Quantitative involvement

### Operationalizing quantitative involvement

Quantitative involvement has been researched extensively in terms of the absolute amount of time parents spend, or tasks they perform, in caring for the child [see @pleckPaternalInvolvementRevised2004]. However, some have argued that the the way childcare tasks are divided is more relevant for children's adjustment than the absolute amount of time invested by each parent [@krollEarlyFatherInvolvement2016; @sabattiniRelationMothersFathers2004; @craigHowMothersFathers2011; @feinbergInternalStructureEcological2003]. Indeed, research has shown that the equity of the childcare distribution affects child development directly [@deutschPaternalParticipationChild2001], and impacts the emotional climate of the interparental relationship [@craigDoesFatherCare2006]. When childcare is unequally distributed, women often perform a disproportionate share of the work. This strains the marital relationship [@shockleyItNotWhat2018], and the resulting stress erodes parenting quality, particularly for parents of difficult children [@sheltonInterparentalConflictNegative2008]. Thus, operationalizing quantitative involvement in terms of relative involvement might capture patterns of greater relevance.

There are also methodological advantages to focusing on the distribution of childcare. Operationalizing quantitative involvement in terms of both parents' absolute involvement incurs a potential confound: Any effect that appears to be due to fathers' greater involvement might, in fact, be due to children receiving greater a greater total amount of parental care [see @deutschPaternalParticipationChild2001]. This risk is exacerbated by the fact that, within families, fathers' absolute involvement is strongly associated with mothers' absolute involvement [@russellCoparentingEarlySchoolage1994]. Operationalizing quantitative involvement in terms of the childcare distribution, and controlling for total joint care, mitigates this confound. Some prior research has attempted to capture the childcare distribution directly, by asking which parent is most responsible for certain tasks, on a continuum from the mother doing everything to the father doing everything [@krollEarlyFatherInvolvement2016]. When quantitative involvement has been measured in terms of absolute involvement, however, an elegant statistical solution can be applied to disentangle the childcare distribution from total joint care. The so-called "latent mean and difference" approach, which is typically used to capture the difference between consecutive longitudinal measurements [@mcardleLatentVariableModeling2009], can be utilized to re-express mothers' and fathers' absolute quantitative involvement into a "latent mean" of involvement, which represents the total joint care provided by both parents, and latent differential involvement, which represents the extent to which one parent is more involved than the other. This approach taps directly into a theoretically relevant construct, and prevents confounding.

A final consideration in the operationalization of quantitative involvement is that the type of tasks parents perform in the context of child care change rapidly as children develop [@wangATUSFathersInvolvement2009]. By the time children are school-aged, their dependence on instrumental care might be relatively minor. When the behaviors constituting quantitative involvement change over time, it is not clear how one might establish measurement invariance on a macro-longitudinal time scale. Consequently, in the present study, we took quantitative involvement at the first measurement wave as an indicator of between-individual differences in quantitative involvement. This means that we cannot investigate within-family dynamics pertaining to change in quantitative involvement.

## Main effects of quantitative involvement

### Effects on qualitative involvement

In many countries, mothers continue to spend more time with their children than fathers [@hofferth28669resident], particularly when it comes to routine care tasks [@craigHowMothersFathers2011]. Fathers' involvement is on the increase, however [@yeungChildrenTimeFathers2001]. The amount of time parents spend with children is likely to predict the quality of parenting behaviors [@sabattiniRelationMothersFathers2004], through practice and experience. Indeed, more involved parents tend to experience greater parenting self-efficacy [@gialloMothersFathersInvolvement2013a]. We might thus hypothesize that greater total joint care should positively predict both parents' parenting quality. This hypothesis is also supported by the notion of parental "buffering", which holds that, when both parents are more involved, they monitor each other's parenting behaviors, thereby ensuring that children are raised well [@mclanahanConsequencesFatherAbsence2014].

The notion that parenting quality improves through practice [@sabattiniRelationMothersFathers2004] might suggests that, when one parent is more involved than the other, the more involved parent would display more desirable qualitative parenting behaviors. However, an argument to the contrary can be made as well. The notion of parental buffering would suggest that both parents' qualitative parenting behaviors might suffer when parents are not equally involved [@mclanahanConsequencesFatherAbsence2014]. Indeed, retrospective reports by 19-year old participants indicate that an unequal childcare distribution tends to go hand in hand with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004]. Greater differential involvement might undermine both parents' involvement quality, through its effect on the marital relationship. Childcare responsibilities are a heavy burden if they come down on a single parent's shoulders [@deutschPaternalParticipationChild2001]. Mothers thus benefit directly when fathers relieve some of this burden [@pleckPaternalInvolvementRevised2004]. Consequently, mothers tend to be more satisfied when fathers are more involved [@pharesFathersMothersInvolvement2009]. Moreover, balanced co-parenting appears to be associated with more desirable qualitative parenting behaviors; at least in terms of parental warmth [@bondsExaminingProcessWhich2007; @barnettInterdependenceParentingMothers2008]. An unfair distribution of childcare, on the other hand, is associated with greater maternal anger and resentment [@rossGenderParenthoodAnger1996]. This strains the interparental relationship, interferes with sensitive parenting, and undermines child wellbeing [@davies2006interparental]. In other words, when care is distributed relatively equally, the entire family system will likely be more harmonious. We might thus expect greater differential involvement to aversely affect both parents' parenting quality, as well as children's emotional adjustment.

### Effects on emotional adjustment

Although the quantity of paternal involvement has received substantial interest in the fields of sociology and gender studies, this literature has focused primarily on its implications for mothers' time use and wellbeing. Consequences for children's emotional outcomes have received less attention to date. Bioecological theories of development indicate that children require a certain amount of care, and who provides this care is, to some extent, flexible [see @bronfenbrennerEcologyHumanDevelopment1979]. This suggests that the total amount of care provided should be positively associated with children's emotional adjustment. Differential involvement, on the other hand, might be negatively linked to emotional adjustment: When children perceive that one of their parents is relatively uninvolved, they likely feel less supported than when both parents are equally invested in their wellbeing [@deutschPaternalParticipationChild2001].

<!--Fathers are as capable as mothers of caring for their children **(Lamb 1997; Yeung et al. 2001)**, and on average, -->

<!-- dit mogelijk vóór de effecten van quant plaatsen? -->

## Interactions between involvement quality and quantity

Interactions between involvement quality and quantity have largely been overlooked in the aforementioned theoretical models of father involvement; perhaps in part because these two dimensions have rarely been jointly investigated [@jaffeeLifeFatherBenefits2003]. Children's emotional adjustment might be differently impacted by paternal control if that father is generally highly involved with the day-to-day minutiae of childrearing, than when he is generally absent. Conversely, children's emotional adjustment might elicit different qualitative parenting behaviors from hands-on fathers, who are aware of the reasons for their children's emotional troubles, than from out-of-touch fathers, who might perceive maladjustment as a problem behavior that should be sanctioned. To capture patterns like these, a model should account for interactions between involvement quality and quantity.

In this respect, theories of father involvement can be complemented by Feinberg's coparenting model [@feinbergInternalStructureEcological2003]. Coparenting refers to the ways parents relate to each other in the role of parent. Although the coparenting literature is typically invoked to describe childrearing in divorced families, it also holds insights for intact families. One dimension of coparenting is involvement quantity, which Feinberg defines as the "division of labor" in childcare. Note that this definition parallels our operationalization of quantitative involvement. The division of childcare labor is proposed to affect children's adjustment and parents' qualitative childrearing behaviors directly, and to interact with the link between these two constructs.

There is some preliminary empirical support for interactions between involvement quality and quantity; for instance, fathers' playful and caring involvement predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@jiaTransactionalRelationsFather2012]. Furthermore, in adolescence, the quantity of paternal education-related structure, an age-appropriate form of involvement, interacted with paternal warmth to predict both academic achievement and behavioral adjustment [@wangParentalInvolvementAfrican2014]. Applying this insight from the coparenting literature, we might thus expect parents' quantitative involvement to be directly associated with their qualitative parenting behaviors and children's emotional adjustment, and to moderate effects between qualitative parenting behaviors and emotional adjustment.

## Sex differences

<!-- pubertal timing and associated changes in hormone levels during adolescence are related to sex-typed attributes, including emotional distress (Brooks-Gunn, Graber, & Paikoff, 1994; Susman, Dorn, & Chrousos, 1991).-->

There is substantial evidence that parents tend to raise boys and girls differently. Prior research has indicated that parents tend to be more controlling with girls than with boys [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and express more negative parenting with boys than with girls [@strausCorporalPunishmentAmerican1999]. Another point to consider is the interplay between parent and child sex. According to social learning theory, the same-sex parent plays a stronger role in a child's development [@busseySocialCognitiveTheory1999], particularly in the socialization of sex-typed behavior, such as display rules for emotions [@brodySocializationGenderDifferences2000]. Thus, we might expect child sex to predict the same-sex parents' qualitative involvement, and associations between parenting and emotional adjustment might be stronger for same-sex parent-child dyads. There is also evidence that parental sensitivity is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a]. These findings suggest that boys tend to receive a "tougher" upbringing than girls. In sum, it is clear that sex differences are a relevant factor in the interplay between parenting and emotional adjustment. Specifically, we might expect to find parents, display more responsiveness, structure, and inductive reasoning with girls than with boys. Boys, on the other hand, might receive more negative parenting. Furthermore, we might expect associations between qualitative parenting behaviors and emotional adjustment to be stronger between children and the same-sex parent.

There is little evidence for sex differences in emotional adjustment in young children, as indicated by a meta-analysis which found no sex differences in the temperament dimension of negative affectivity, based primarily on infant and toddler samples [@else-questGenderDifferencesTemperament2006a]. Evidence for such sex differences is sometimes found in older children [@silkAdolescentsEmotionRegulation2003], but these might appear and intensify in puberty [@hillIntensificationGenderRelatedRole1983]. We consequently did not expect to find sex differences in children's average levels of emotional adjustment.

## The present study

The present study aimed to make several contributions to the literature on fathers' role in their children's emotional adjustment development. First, we included both mothers and fathers, in order to tease apart their common and unique roles. Second, our operationalization of involvement incorporated both qualitative involvement, in terms of specific parenting behaviors, and quantitative involvement, operationalized as total joint childcare and differential involvement. Third, we employed recent methodological advances to disentangle stable between-family differences from dynamic within-family predictive processes, in order to more accurately assess the direction of effects between qualitative parenting behaviors and emotional adjustment. Fourth, we examined interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work.

Based on our review of the literature, we expected specific dimensions of qualitative involvement to be associated with emotional adjustment at both the within- and between-family level. The literature suggests that warmth, consistency, and inductive reasoning should be positively associated, whereas negative parenting should be negatively associated, with emotional adjustment. With regard to child effects, we might expect some of these paths to be reversed: Positive adjustment might lead to diminished inductive reasoning and parental anger. We further hypothesized that quantitative involvement would predict children's adjustment directly: Greater joint total care should predict better emotional adjustment, whereas a more skewed distribution of care should predict poorer emotional adjustment. Greater joint care should also predict both parents' qualitative involvement, in the sense of lower parental anger, and greater consistency, inductive reasoning, and warmth. A more skewed distribution of care, on the other hand, should show these associations only for the more involved parent. Importantly, we also hypothesized that involvement quality and quantity would interact, such that a more skewed distribution of quantitative involvement would interfere with the effects of qualitative parenting behaviors on children, and parents' adaptive responses to children's maladjustment. In other words, we expected that an unfair childcare distribution would predict suboptimal parenting, and parental responses to maladjustment, due to the strain such a distribution places upon the entire family system.  Finally, we accounted for potential effects of child sex.

# Methods

## Participants

We employed data from the`r masked("K-cohort")` of the `r masked("Longitudinal Study of Australian Children (LSAC)")`, waves `r masked("2-5")`, which were collected by means of biannual home visits, and self-report mail questionnaires, from 2006-2012. `r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population. Complete details regarding the study design, sample, and procedure are available online at `r masked("http://data.growingupinaustralia.gov.au/")`.

Our target population consisted of a subsample of 1884 native `r masked("Australian")`, non-minority, intact, biological families. At the first wave, children were approximately 6 years old, $M = 6.28$, $SD = .45$, and 12 at the final wave ($M = 12.42$, $SD = .49$); on agerage, mothers were 34.89 ($SD = 4.71$), and fathers 36.94 ($SD = 5.48$). Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$; 669 father-only, 41 mother-only). The modal annual income for fathers was between \$26,000-\$51,999 (N = 663), and for mothers, \$25,999 or less per year (N = 1152).

## Measures

### Qualitative involvement.

Parents completed measures of several qualitative parenting dimensions. Responsiveness and inductive reasoning were measured using the warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire [CRQ, @patersonAssociationBehaviouralAdjustment2001]. Structure and negative parenting were measured using the consistency and anger scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000]. For these instruments, parents reported on their own behavior using five-point scales, ranging from "never or almost never" to "always or almost always".

### Quantitative involvement.

Measures of quantitative involvement varied from wave to wave, due to children's changing care needs [@wangATUSFathersInvolvement2009]. We used reports at T1 as an indicator of between-family differences in quantitative involvement. Both parents rated their involvement in four different child care activities on a five-point scale ranging from "once a day or more" to "not at all" [e.g., "In the past month how often did you help this child brush his/her teeth?"; @baxterFatheringAustraliaCouple2011].

### Emotional adjustment.

Emotional adjusmtment was measured using the reverse-coded emotional distress subscale of the Strengths and Difficulties Questionnaire, reported by teachers on a three-point scale from "not true" to "certainly true" [e.g., "Often unhappy, downhearted or tearful"; @goodmanPsychometricPropertiesStrengths2001].

## Scale construction

We calculated mean scores for all of the study variables. Scale descriptive statistics are provided in Supplementary Table 1, and scale intercorrelations are provided in Supplementary Table 2. To establish that metric measurement invariance could be imposed across measurement waves, and across mothers and fathers, we estimated structurally invariant CFA's for all of the study variables, and then imposed constraints on the factor loadings across time and reporter. Chi-square difference tests did indicate significantly worse fit for metric invariance models than for structural invariance models, (all $ps \leq$ .03$). However, chi-square difference tests are overpowered in large samples. The difference in fit between the structural invariance models and metric invariance models was smaller than .01 according to conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3). This indicates that it is defensible to assume that the questionnaires measure the same constructs across time and reporter.

## Strategy of analyses

We used multilevel, autoregressive structural equation modelling to address our research questions [@asparouhovDynamicStructuralEquation2018]. In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean). In the multilevel model, these stable individual differences are represented by random intercepts at the between-persons level. Fluctuations from these random intercepts are analyzed at the within level. Time-lagged copies of all within-level variables are created, in order to estimate a cross-lagged autoregressive model. The coefficients of this cross-lagged autoregressive model are free to vary between individuals, and can thus be predicted by between level moderators, such as quantitative involvement and child sex.

This approach has several advantages over the classic fixed-time cross-lagged panel model. First, controlling for stable between-individual and between-family differences allows us to obtain more accurate estimates of mothers' and fathers' unique roles in child development [@hamakerCritiqueCrosslaggedPanel2015; @berryPracticalInterpretabilityCrosslagged2017]: Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts. Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects. In this way, we included total joint childcare, the childcare distribution, and child sex as moderators.

All analyses were conducted in Mplus 8.0, controlled from R using MplusAutomation. In Mplus, this type of multilevel autoregressive model can currently only be estimated using Bayesian methods. We used Gibbs sampling to estimate the model, with two chains of 50000 iterations each. We used mildly informative priors for the variance/covariance matrices, and regularizing priors centered on zero for all regression coefficients. We assessed convergence through visual inspections of the chains and posterior distributions. We report the posterior median, and a 95% highest posterior density interval (HPDI). This interval is the "narrowest" possible interval of the posterior distribution that contains 95% of samples from the posterior distribution. It represents the most likely population values of the parameter in question, given the model, priors, and data. Results are best interpreted by examining the effect size (standardized regression coefficients and correlations), and the width of the HPDI. However, for convenience's sake, we focus on interpreting results whose HPDI did not include zero. Significance is not a Bayesian concept, but we use the shorthand "significant" to refer to estimates for which less than 5% of the posterior distribution had the opposite sign from the estimate; i.e., for a positive estimate, this is the proportion of the posterior below zero. A complete overview of the standardized and unstandardized model parameters is provided in Supplementary Table 4.

# Results

<!--structure notes
1) mothers and fathers, in order to tease apart their common and unique roles. 
2) incorporated quantitative involvement, in terms of total joint childcare and the childcare distribution, and qualitative involvement
3) interactions between qualitative and quantitative involvement
4) assess the direction of effects between parenting behaviors and emotional adjustment.

HYP:
3) Qualitative parenting associated with emotional adjustment at both the within- and between-family level
   * responsiveness, structure, and reasoning positively associated, negative parenting negatively associated
   * Correlations, and regressions 
     * Parent effects
   * Child effects, some paths reversed: Maladjustment should elicit decreased responsiveness, but increased structure, reasoning, and negative parenting. 
1) Interactions between involvement quality and quantity
   * skewed distribution of quantitative involvement would interfere with the effects of qualitative parenting behaviors on children (parenting effects), and parents' adaptive responses to children's maladjustment (child effects).
   * unfair childcare distribution would predict suboptimal parenting
2) Direct effects of quantitative involvement on adjustment and parenting
   2.a) Adjustment
   * Greater joint total care should predict better emotional adjustment
   * a more skewed distribution of care should predict poorer emotional adjustment
   2.b) Greater joint care should also predict both parents' qualitative involvement
   * greater responsiveness and structure, and lower negative parenting.
   * A more skewed distribution of care, on the other hand, should show these associations only for the more involved parent.

4) child sex
-->

## Associations between qualitative parenting dimensions and emotional adjustment

We hypothesized that, at both the within- and between-family levels, anger would be negatively associated, and consistency, inductive reasoning, and warmth would be positively associated with emotional adjustment.

### Between-family level correlations

At the between-family level, we examined correlations between the random intercepts of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabbetweencor)). The random intercept of each family is based on their average level across all waves of the study. Correlations between random intercepts are estimated across all families in the sample. Results provided mixed support for our hypotheses: Contrary to hypotheses, we found no significant correlations between maternal and paternal warmth and children's emotional adjustment. In support of hypotheses, we did find significant positive correlations between maternal and paternal consistency, and children's emotional adjustment. Contrary to hypotheses, we found significant negative correlations between maternal and paternal inductive reasoning, and children's emotional adjustment. This means that, in families where parents use more inductive reasoning, children displayed poorer emotional adjustment. Finally, as hypothesized, we found significant between-family level negative correlations between maternal and paternal anger, and emotional adjustment.

```{r tabbetweencor, results = "asis"}
cor_tables <- lapply(results, corTable)

table_between_cor <- lapply(cor_tables, `[[`, "Within.Std.Averaged.Over.Clusters")
table_between_cor <- order_by_name(table_between_cor)
table_between_cor <- do.call(cbind, table_between_cor)
table_between_cor <- cbind(paste0(1:3, ". ", rename_pars(rownames(table_between_cor))), table_between_cor)
table_between_cor <- as.matrix(table_between_cor)
colnames(table_between_cor) <- c("Parameter", rep(paste0(1:3, "."), 4))
rownames(table_between_cor) <- NULL
table_between_cor <- gsub("\\*+", "*", table_between_cor)
table_between_cor <- table_between_cor[, -seq(2, 11, by = 3)]
write.csv(table_between_cor, "table_between_cor.csv", row.names = FALSE)
apa_table(table_between_cor, caption = 
            "Between-families level correlations", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.", col_spanners = colspan_fun(table_between_cor, skip = 1)
          )

```

<!--(Discussion: Although inductive reasoning is considered to be a disciplinary behavior that stimulates children's positive development (better ref here), the fact that it is a disciplinary behavior might explain the negative association with emotional adjustment.).-->

### Within-family correlations

We also examined correlations between qualitative parenting behaviors and emotional adjustment at the within-family level (Table \@ref(tab:tabwithincor)). Within-family correlations are based on residuals of the study variables, across all waves and families. In support of hypotheses, we found significant positive within-wave associations between emotional adjustment and fathers' and mothers' warmth. Notably, these associations were nearly twice as strong for fathers' warmth, as compared to mothers' warmth. In partial support of hypotheses, we did find significant positive associations of adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning. Finally, contrary to our expectations, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger.

```{r tabwithincor, results = "asis"}
table_within_cor <- lapply(cor_tables, function(x){x[["Between"]][3:5, 3:5]})
table_within_cor <- order_by_name(table_within_cor)
table_within_cor <- do.call(cbind, table_within_cor)
table_within_cor <- cbind(paste0(1:3, ". ", rename_pars(rownames(table_within_cor))), table_within_cor)
table_within_cor <- as.matrix(table_within_cor)
colnames(table_within_cor) <- c("Parameter", rep(paste0(1:3, "."), 4))
rownames(table_within_cor) <- NULL
table_within_cor <- gsub("\\*+", "*", table_within_cor)
table_within_cor <- table_within_cor[, -seq(2, 11, by = 3)]
write.csv(table_within_cor, "table_within_cor.csv", row.names = FALSE)
apa_table(table_within_cor, caption = 
             "Within-family residual correlations", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.", col_spanners = colspan_fun(table_within_cor, skip = 1)
          )
```

### Within-family predictive effects

At the within-family level, we examined over-time cross-lagged effects between qualitative parenting and emotional adjustment. The intercepts of these random effects reflect the average cross-lagged effect across families. The variances indicate whether the effect differs in strength (and sign) between families.

#### Parenting effects

None of the intercepts of within-families parenting effects differed significantly from zero, excepting the effect of maternal warmth on children's emotional adjustment (Table \@ref(tab:tabparenteff)). Thus, on average across families, mothers' warmth had a positive significant predictive effect on children's emotional adjustment. The variances of all parenting effects were significant, however, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families. Moreover, these variances were notably larger for fathers than for mothers, indicating that fathers have a much more variable effect on their children through specific parenting behaviors than mothers do.

```{r tabparenteff, results = "asis"}
tabparenteff <- rbind(tab_std_sections$`Intercepts of parenting effects`, 
                      tab_std_sections$`Between-family level residual variances`[c(1, 5:6),])

tabparenteff$Parameter <- gsub("^\\(.+on (.+)\\)$", "\\1", tabparenteff$Parameter)
tabparenteff$Parameter[c(1,4)] <- c("Random intercepts of parenting effects", "Variance of parenting effects")
reg_table(tabparenteff, caption = "Parenting effects on emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

#### Child effects

In line with hypotheses, the intercept of the cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabchildeff)). Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more anger and inductive reasoning, but also more warmth, than usual in the next year. Contrary to hypotheses, the intercept of the effect of emotional adjustment on maternal consistency was non-significant. As hypothesized, the intercept for maternal inductive reasoning was positive. Finally, counter to hypotheses, the intercept for maternal anger was positive, not negative.

In partial support of hypotheses, the intercept for the child effect of adjustment on paternal warmth was positive and significant. None of the other intercepts for child effects on paternal parenting were significant.

The variances of all child effects were significant, indicating that the effect of children's emotional adjustment on their parents' childrearing behaviors differs between families.

```{r tabchildeff, results = "asis"}
tabchildeff <- rbind(tab_std_sections$`Intercepts of child effects`,
                      tab_std_sections$`Between-family level residual variances`[c(1, 7:8),])

tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
tabchildeff$Parameter[c(1,4)] <- c("Random intercepts of child effects", "Variance of child effects")
reg_table(tabchildeff, caption = "Child effects on qualitative parenting", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

## Associations between quantitative involvement and parenting and adjustment

We examined regression coefficients from total joint involvement and relative involvement to the random intercepts of emotional adjustment and qualitative parenting behaviors (Table \@ref(tab:tabinvolvement)). With regard to parenting, we found partial support for our hypotheses: For both parents, joint involvement positively predicted greater warmth, consistency, and inductive reasoning. Notably, the association between total joint involvement and warmth was almost double in size for fathers, as compared to mothers. We did not find the hypothesized negative effects of joint total involvement on parental anger.

Counter to our hypothesis, we found that greater joint care was negatively associated with emotional adjustment. Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment. With regard to the distribution of childcare, however, our hypothesis that children's emotional adjustment would be poorer in families where mothers were relatively more involved than fathers was supported.

With regard to relative involvement, we found partial support for our hypothesis that more skewed relative involvement should predict greater warmth, consistency, and inductive reasoning, and reduced anger, only for the more involved parent. For mothers, the only significant association we found was that, in line with this hypothesis, mothers displayed more warmth when they were relatively more involved than fathers. For fathers, we found that, in line with this hypothesis, fathers displayed less warmth, consistency, and inductive reasoning, when mothers were relatively more involved than fathers.

```{r tabinvolvement, results = "asis"}
tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
tabinv$Parameter[c(1,4)] <- c("Effects of joint quantitative involvement on:",
                              "Effects of relative quantitative involvement on:")
tabinv$Parameter <- gsub("^(.+) on.*$", "\\1", tabinv$Parameter)
reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

## Interactions between involvement quality and quantity

To examine interactions between involvement quality and quantity, we tested whether between-family differences in quantitative involvement predicted the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvementmod)).

```{r tabinvolvementmod, results = "asis"}
tabinvmod <- rbind(
  tab_std_sections$`Moderating effect of quantitative involvement on parenting effects`[c(1:3, 1, 4:5),],
  tab_std_sections$`Moderating effect of quantitative involvement on child effects`[c(1:3, 1, 4:5),])
tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on parenting effects of:",
                              "Moderating effects of relative involvement on parenting effects of:",
                              "Moderating effects of joint involvement on child effects on:",
                              "Moderating effects of relative involvement on child effects on:")
reg_table(tabinvmod, caption = "Moderating effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

### Moderation of parenting effects

With regard to the moderation of parenting effects by quantitative involvement, the only significant effects we found were for maternal warmth. Specifically, the within-families predictive effect of maternal warmth on emotional adjustment was negatively when both parents' total joint involvement was higher. This same effect was positively attenuated, however, if mothers' involvement was relative higher than fathers' involvement. We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.

### Moderation of child effects

With regard to the moderation of child effects by quantitative involvement, we found that the effect of adjustment on maternal warmth was negatively attenuated when parents' joint involvement was higher, and also when mothers' relative involvement was greater compared to fathers. We found no effects for consistency. For inductive reasoning, we found that that the cross-lagged effect of adjustment on maternal inductive reasoning was negatively attenuated when both parents' joint involvement was higher. In other words, when both parents were  more involved, the cross-lagged effect of children's adjustment on maternal discipline was more negative. Adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers. Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment more strongly predicted decreased paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning. We found no moderating effects for anger.

<!--	(Discussion: i.e., if the child is well-adjusted, that leads to less maternal maternal warmth, but only when both parents are more involved. Paradoxical, but keep in mind that maternal and paternal warmth are higher in families with greater joint involvement. So perhaps this means that mothers of maladjusted children, in families where joint involvement is low, coddle their children.)-->

## Effects of child sex

To explore sex differences, we included child sex as a predictor of the random intercepts for emotional adjustment and qualitative parenting behaviors, and of the within-family cross-lagged effects.

With regard to sex differences in the random intercepts, we found no sex differences in emotional adjustment (Table \@ref(tab:tabsex)). For qualitative parenting behaviors, we found that both mothers and fathers displayed less warmth, and more inductive reasoning and anger, towards sons as compared to daughters. No significant differences were found for maternal warmth, and for both parents' consistency.

The only cross-lagged effects significantly moderated by child sex were child effects of emotional adjustment on maternal and paternal inductive reasoning. Specifically, emotional adjustment was more negatively predictive of parents' inductive reasoning for boys, as compared to girls. In other words, if children display better adjustment than usual in one year, their parents decrease inductive reasoning more strongly in the next year when children are male.

```{r tabsex, results = "asis"}
tabsex <- rbind(
  tab_std_sections$`Sex differences in parenting and adjustment`,
  tab_std_sections$`Sex differences in cross-lagged effects`[c(1:3, 1, 4:5),])
tabsex$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabsex$Parameter)
tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
tabsex$Parameter[c(1,5,8)] <- c("Sex differences in parenting and adjustment",
                              "Sex differences in parenting effects of:",
                              "Sex differences in parenting effects on:")
reg_table(tabsex, caption = "Sex differences", note = "Sex is a binary predictor (0 = female, 1 = male). Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

# Discussion

We set out to disentangle mothers' versus fathers' unique roles in children's emotional adjustment development. In doing so, we incorporated both qualitative and quantitative measures of parental involvement, and examined interactions between these two aspects of involvement, which have historically been examined mostly in isolation. To this end, we used innovative methods to include between-family differences in quantitative involvement, operationalized as total joint care and the relative childcare distribution, as predictors of within-family cross-lagged effects between qualitative involvement behaviors and emotional adjustment.

## Associations between qualitative parenting and emotional adjustment

We hypothesized that, at both the within- and between-family levels, warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment. Breaking down the results by parenting behavior, we found mixed evidence for these hypothesized patterns.

We found partial, but unequivocal, support for the hypothesized positive associations between warmth and emotional adjustment. These links were absent at the between-family level, but abundant at the within-family level. Specifically, we found within-family correlations for both parents' warmth; an average prospective effect of maternal warmth on children's emotional adjustment, and child effects of adjustment on both parents' warmth. In other words, well-adjusted children evoke more parental warmth. This was the only socialization effect that was significant on average, which suggests that it occurs in most families. However, it is important to note that *all* within-family cross-lagged effects had significant variance components, highlighting the fact that the findings discussed here only reflect the average across all families, and that dynamic reciprocal processes vary from one family to the next.

We also found partial support for the hypothesized positive associations of parental consistency with emotional adjustment. These associations occurred primarily at the between-family level, where both maternal and paternal consistency were positively correlated with children's emotional adjustment. This means that, in families where parents tend to display high consistency, children tend to be more emotionally well-adjusted. In terms of within-family correlations, we found the predicted links for mothers', but not fathers', consistency. On average, we found no support for within-family predictive links between consistency and adjustment. It is interesting that we found only between-family links and no over-time predictive effects. There are several reasons why between-family links might be observed: For instance, these can reflect the accumulated effect of within-family causal effects at an earlier age. However, our study included children from the age of six, and - although it is possible that consistency did, on average, affect children's emotional development prior to the preschool years, it seems unlikely that this effect would have dissipated entirely by the time our study commenced. Alternatively, associations at the between-family level can reflect stable differences between families that impact both constructs; for example, it has been proposed that the economic hardship associated with low SES diminishes parents' capacity for consistent parenting, while at the same time, undermining children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990]. The fact that we observed no within-family predictive effects, despite including children from a very young age, might suggest that previously found associations between consistency and emotional adjustment might be less attributable to causal processes than prior research had concluded, based on methods that did not control for stable between-family differences [@brandPerceivedParentalInconsistency1990].

With regard to inductive reasoning, we found mixed support for the hypothesized positive associations with emotional adjustment: Associations were opposite in sign at the between- versus within-family levels. At the between-family level, parents who engaged in more inductive reasoning had children with poorer emotional adjustment. This apparently paradoxical finding might be explained by considering that inductive reasoning is a disciplinary behavior, although a constructive one that stimulates children's internalization of rules and expectations [@grusecImpactParentalDiscipline1994]. Parents' reports of engaging in inductive reasoning might thus be confounded with how often they discipline their children. Disciplining children might result in, be precipitated by, or share a common cause with emotional maladjustment. Paradoxically, at the within-family level, emotional adjustment was positively correlated with paternal inductive reasoning, and emotional adjustment had a positive prospective effect on maternal inductive reasoning. In other words, increased maladjustment goes hand in hand with increased inductive reasoning. The fact that between-family differences and within-family fluctuations showed opposing patterns of associations with adjustment might suggest that between-family differences in inductive reasoning tap more strongly into the amount of discipline parents engage in, whereas the within-family fluctuations might more strongly reflect changes in the way parents discipline their children.

With regard to parental anger, we found mixed evidence for the predicted negative associations with emotional adjustment. We primarily found support for these associations at the between-family level (i.e., in terms of the random intercept correlations). This means that, in families where mothers or fathers display low anger, children tend to be more emotionally well-adjusted. Contrary to our expectations, however, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger. Similarly, the average within-family predictive effects of parental anger on adjustment were non-significant. This suggests that the brunt of negative associations between anger and adjustment are explained by differences between families. One potential explanation for the absence of within-family associations might be a ceiling effect; it might be the case that, once the stable level of parental anger in a family is high enough to negatively impact children's emotional adjustment, over-time fluctuations of that anger no longer make much of a difference. Another explanation might be that any further effects of fluctuations in anger play out on a different time scale than the one captured by our study; for example, if such effects primarily play out immediately, in dyadic interactions between parents and children. Finally, we found that, on average, emotional adjustment showed a paradoxical positive predictive effect on maternal anger. This means that, if children demonstrated better emotional adjustment than usual in one year, their mothers tended to display more anger than usual in the next year. The literature on expectation-based loss aversion offers a possible explanation for this puzzling finding [@koszegiModelReferenceDependentPreferences]. Note that, on average, emotional adjustment showed no-to-negative stability coefficients in all models. That means that, when children demonstrated better emotional adjustment than usual in one year, this was more likely than not to be followed by a year with worse adjustment. Mothers, on the other hand, might expect their children's emotional adjustment to be relatively stable. When a child falls short of the mother's expectations for stable emotional adjustment, the perceived "lost utility" resulting from having to deal with a child that is less well-adjusted than expected might render a mother more prone to experiencing anger towards the child.

## Quantitative involvement and qualitative parenting behaviors

Associations between quantitative involvement and qualitative parenting behaviors were in line with the notion that greater involvement is associated with more desireable parenting behaviors [@sabattiniRelationMothersFathers2004]. When both parents were more involved, they displayed greater warmth, consistency, and inductive reasoning. Fathers, in particular, reported substantially greater parental warmth when both parents were more involved. When childcare was more unequally distributed, with mothers performing more care tasks than fathers, we found that mothers displayed more warmth, and fathers displayed less warmth, consistency, and inductive reasoning. The fact that mothers displayed more warmth when fathers were relatively less involved might reflect an attempt to compensate for his absence [@shinnFatherAbsenceChildren1978]. In sum, these results suggest that fathers engage in less desirable qualitative parenting behaviors, in families where childcare is unequally distributed [@sabattiniRelationMothersFathers2004; @craigHowMothersFathers2011].

## Quantitative involvement and emotional adjustment

The present study operationalized quantitative involvement in terms of total joint care provided by parents, and the childcare distribution, which captures differences between mothers and fathers. This operationalization addresses a criticism raised in the literature, that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001]. For total joint care, we found that, counter to our hypothesis, joint care was *negatively* associated with between-family differences in children's emotional adjustment. One potential explanation of this finding is suggested by self-determination: Perhaps excessive levels of involvement undermine children fundamental needs for autonomy and a sense of competence, thereby leading to increased emotional maladjustment [@deciFacilitatingOptimalMotivation2008]. Research on "helicopter parenting" provides some support for this interpretation, as psychological wellbeing has been found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014]. It should be noted, however, that this literature is mostly cross-sectional, and focused on adolescents. Another potential explanation is that the amount of quantitative care provided is not only driven by parents' motivation and ability to be involved in their children's lives, but also by children's objective needs. Factors related to children's objective needs - such as developmental delay, or atypical development - might thus drive an increase in parents' quantitative involvement, and might also be associated with children's emotional maladjustment, thereby resulting in a negative correlation between the two.

In line with predictions derived from coparenting theory, we found that children's emotional adjustment was lower in families where mothers were more involved than fathers - even after controlling for the total amount of childcare provided by parents. Moreover, mothers were more involved than fathers, on average, as indicated by the positive significant mean of the latent variable capturing differential involvement. This suggests that the majority of children experienced the disadvantage in emotional adjustment associated with differential involvement. Although substantial empirical work has indicated that the marital relationship suffers when childcare is unequally distributed [@shockleyItNotWhat2018], our research is one of few to examine the implications for children's adjustment [see also @deutschPaternalParticipationChild2001]. Of course, causality cannot be inferred from correlational findings at the between-family level. Future research should examine potential pathways, such as whether fathers shy away from childcare responsibilities when children display poor emotional adjustment, or whether children's adjustment suffers when they feel like both parents are not equally invested in their wellbeing, or whether differential involvement strains the marital relationship, and this strain in turn spills over and affects children's emotional adjustment.

## Interactions between qualitative and quantitative involvement

We found limited evidence that within-family parenting effects - cross-lagged effects from qualitative involvement behaviors on emotional adjustment - were moderated by quantitative involvement. One such effect, which was in line with predictions, is that the effect of maternal warmth on adjustment was contingent on both total joint care, and the relative childcare distribution. Mothers' warmth predicted children's adjustment more strongly if mothers were relatively more involved than fathers. Conversely, mothers' warmth was less strong of a predictor of adjustment development when both parents were more involved. Together, these findings suggest that having a warm mother is of greater developmental importance with regard to children's emotional adjustment if fathers are less involved with the day-to-day minutiae of childcare, and is less essential when both parents are highly involved. This suggests a compensatory effect, whereby mothers' warmth can make up for any disadvantages incurred by having a relatively uninvolved father [@shinnFatherAbsenceChildren1978].

Interactions between qualitative and quantitative involvement were more abundant in the context of child effects on parents; i.e., the way parents respond to children's emotional adjustment or maladjustment. Specifically, we found that when mothers were relatively more involved than fathers, fathers decreased their inductive reasoning more strongly in response to children's positive adjustment, and increased it more strongly in response to maladjustment. Considering that inductive reasoning is a form of discipline, this finding suggests that, when mothers are relatively more involved in care tasks, fathers take more of a role as disciplinarian.

<!-- the domains of discipline and play tend to be the areas in which fathers feel most comfortable (Parke 2000).
 Mothers’ satisfaction with the division of labor within the household was consistently related to greater levels of paternal responsibility in all four domains (school work, discipline, daily care, fun activities), but fathers’ satisfaction was only related to greater paternal involvement in two of the four domains (discipline and fun activities). 
-->
This finding is complemented by the negative effect of total joint involvement on the link from emotional adjustment to maternal inductive reasoning, which suggests that mothers also have more of a role as disciplinarian in families where both parents are highly involved. We also found that emotional adjustment became less predictive of maternal warmth when parents' joint involvement was higher, and when mothers' relative involvement was greater compared to fathers. This finding can be interpreted in the context of the sizable average effect of adjustment on maternal warmth. On average, increases in emotional adjustment predicted substantial increases in maternal warmth. In families where mothers' involvement was higher - either in terms of joint, or relative involvement - maternal warmth was thus less contingent on children's adjustment.

## Differences between fathers and mothers

Overall, we found somewhat more evidence for associations of emotional adjustment with mothers' qualitative parenting, than with fathers' parenting, and we found some evidence for differences between mothers' and fathers' roles in emotional adjustment development. These differences all played out at the within-family level, however. Stable between-family differences in either parent's qualitative parenting were all similarly associated with children's adjustment.

At the within-family level we found stronger associations with emotional adjustment for maternal consistency, and for paternal reasoning and warmth. As mentioned before, mothers tend to spend more time with their children than fathers; also in our sample. This might help explain the stronger associations for maternal consistency - a behavior which implies, by definition, repeated interactions with the child. For fathers, associations were stronger for inductive reasoning and warmth - behaviors which can be expressed in one-off interactions, such as when disciplining a child (inductive reasoning), or during play time (warmth). Moreover, the fact that the association between inductive reasoning and emotional adjustment was stronger for fathers than mothers is in line with the classic view of the father as disciplinarian [@ferrariImpactCultureChild2002].

With regard to parenting effects, the only difference between mothers and fathers we found was for the predictive effect of warmth on emotional adjustment, which was found only for mothers. This is in line with prior research, which suggested that warmth is an important socialization mechanism for mothers [@vanlissaRoleFathersMothers2018a].

With regard to child effects on parenting, we found substantial evidence that mothers' parenting was more contingent on children's emotional adjustment than fathers' parenting. Specifically, mothers' anger, inductive reasoning, and warmth were predicted by adjustment, whereas these effects were absent for fathers.

<!-- For anger, the one difference between mothers and fathers was in Within-family predictive effects suggested that, over-time changes in mothers' anger were more contingent on children's emotional adjustment than fathers' anger. We explained this finding in the context of expectation-based loss aversion, resulting from children falling short of mothers' expectations for positive adjustment. The fact that this effect was present for mothers, but not for fathers, might be explained by the fact that mothers typically spend more time with the child [REF] - as evidenced in our study by the significant and positive mean for the latent difference between maternal and paternal involvement. A parent who spends more time with the child might develop stronger expectations for the child's future behavior. -->

## Direction of effects

Historically, the field of developmental psychology focused primarily on how parents "shape", or cultivate, their children's development. Nowadays it is commonly acknowledged that parents and children exert mutual influences. However, commonly used developmental methods have been criticized for failing to properly reveal the direction of effects: When stable between-family differences are not accounted for, estimates of cross-lagged effects can be substantially biased [@hamakerCritiqueCrosslaggedPanel2015]. One recent study examined the interplay between qualitative parenting behaviors and emotion regulation development in adolescence, and found that child effects on parents substantially outweighed parent effects on children [@vanlissaRoleFathersMothers2018a]. The authors speculated that this finding might be due to the relatively egalitarianism of parent-child relationships in adolescence, as compared to earlier childhood. The present study casts doubt on this interpretation, as we found that child effects outweighed parenting effects, even at a young age. Although still preliminary, the prominence of child effects across two large longitudinal studies, conducted on samples from different continents and at different age ranges, invites the question: Are parents really shaping children's development, or are they, instead, responding to it? An extreme interpretation of the findings from our study and this preceding work might be that parents' influence is limited to "setting the stage" for children's development, through their stable differences from other parents, and from there on, are merely adjusting to children's developmental needs. This extreme view can be nuanced by taking into consideration that not all causal effects are evident from the interplay between within-individual fluctuations. For example, although the genetic makeup that is passed from parents to children is stable, the expression of some genes emerges over time. Consequently, stable parent characteristics can cause change in children. Future research might examine alternative theoretical conceptualizations of parental influences in child development. Another point to consider is that parent effects on children might play out on a different time scale than is typically captured in macro-longitudinal studies. If parents shape their children through specific behaviors in dyadic interactions, which are not captured by annual reflective self-reports, then we might fail to capture parenting effects. Future research might test this conjecture by applying DSEM methods to dyadic interaction data.

<!--To what extent these findings should be interpreted as "child-driven" effects, versus "parents adjusting to children's developmental needs".-->

<!--. This suggests that reciprocal dynamic processes might be less prominent than previous research suggested. Instead, third variables, such as socio-economic status, or shared genetic factors, might play a larger role - another outstanding issue on which further research is required (REF Flouri). -->

## Sex differences

In line with prior research, we found that, on average, parents raise boys and girls differently. Specifically, we found that mothers and fathers both displayed more angry parenting and inductive reasoning with sons than with daughters, and that fathers displayed less warmth towards sons. These findings suggest that parents are stricter towards their sons. Parenting effects on children were not moderated by child sex, suggesting that boys and girls are similarly affected by qualitative parenting behaviors. Child effects on parenting, however, were affected: Emotional adjustment was more negatively predictive of both parents' inductive reasoning when children were boys. In other words, parents relax their verbal discipline more in response to positive emotional adjustment when children are male, than when they are female. When viewed in the context of the finding that parents were generally stricter with boys, these findings suggest that some of this greater strictness is conditional on sons demonstrating positive adjustment.

## Strengths and Limitations

This study had several important strengths, which allow us to make unique contributions to the literature. First, mothers and fathers both completed identical parenting measures, which allowed us to disentangle and compare their unique roles in children's emotional adjustment development. Second, we incorporated measures of quantitative involvement, in terms of the amount of time spent in childcare, as well as qualitative involvement, in terms of specific parenting behaviors. This allowed us to provide a more nuanced overview of the way parental involvement is linked with emotional adjustment development, including how the distribution of quantitative childcare relates to parenting quality and emotional adjustment, and moderates links between these two constructs. Additionally, although some prior studies have also incorporated both quality and quantity of parenting, this work has typically included the absolute amounts of maternal and paternal involvement. A limitation of this approach is that it is not possible to distinguish the effect the childcare distribution from the effect of the total amount of care children receive [see @deutschPaternalParticipationChild2001]. We addressed this limitation by re-expressing the variance of maternal and paternal quantitative involvement into joint total involvement, and differential involvement. Third, we used innovative methods to examine the extent to which parenting and adjustment are linked at the between- versus within-family level. Compared to methods that do not distinguish between these levels, our approach provides more accurate assessments of dynamic reciprocal processes within the family. Our findings contributed to a growing literature that suggests that stable between-family differences might play a larger role than was previously thought, even in early childhood. Confidence in the results is bolstered by the fact that we used multi-informant reports: Each parent reported on their own parenting, and teachers reported on children's emotional adjustment. This avoids common method bias and single reporter bias. Furthermore, teachers' reports should be relatively unbiased, because they are likely to be "blind" to parents' childrearing behavior, and have no stake in the parent-child relationship.

One potential limitation of the present study is that teachers' reports of emotional adjustment might be more "noisy" than self- or parent-reports. A teacher is responsible for many pupils, and might not be aware of each pupil's emotional adjustment. Moreover, although the instrument we used assessed *overt* symptoms of emotional maladjustment, people are known to be poor judges of others' emotional states [@ickesEmpathicAccuracy1997]. Consequently, it is important to consider the fact that our effect sizes might be attenuated by noisy measures of emotional adjustment. Teachers' reports do have clear benefits compared to alternatives: Young children might lack the ability to report accurately on their own emotional adjustment, and using clinical assessments might be unfeasible for large-scale panel studies. Another limitation to consider is the fact that we operationalized qualitative parenting in terms of specific parenting behaviors, whereas others have focused on parenting *styles* composed of specific combinations of these behaviors [@maccobySocializationContextFamily1983]. We focused on specific parenting behaviors for two main reasons. First, although aggregating specific parenting behaviors into parenting styles is theoretically defensible, it also amounts to data reduction. This would prevent a fine-grained investigation of differences between mothers' and fathers' unique roles in childrearing. Moreover, the methods we used to disentangle stable between-family differences from within-family processes are not suitable for categorical variables, such as parenting style. Still, we have to acknowledge the possibility that the constellation of parenting behaviors is more important than each dimension in isolation. Finally, one potential limitation is that we only included quantitative involvement in the first year of the study. This was done for pragmatic reasons: Children's developmental needs change rapidly, especially in the first years of life, and consequently, it would not be possible to establish measurement invariance for quantitative involvement from age 6 to 12. For instance, whereas 6-year olds might need help brushing their teeth, 12-year olds typically do not. Our approach assumes that a "snapshot" of quantitative involvement at age 6 is representative for parents' general level of involvement. Consequently, we are unable to assess whether within-person fluctuations in qualitative and quantitative involvement interact. Future research intending to examine such interactions at the within-family level might instead choose to focus on a more restricted age range, so measurement invariance for quantitative involvement can be assured.

## Conclusions

Our findings suggest that, in families where mothers were more involved than fathers, there was greater differentiation between the roles mothers' versus fathers' qualitative parenting played in children's emotional adjustment development. In such families, maternal warmth was more strongly predictive of, and less dependent on, children's emotional adjustment, and fathers' inductive reasoning - a disciplinary strategy - was more contingent on children's emotional adjustment. This suggests that, when differential involvement was higher, parents adopted more differentiated roles: Mothers' role more strongly resembled that of the classic warm "attachment figure", and fathers that of an authority figure. When both parents were more involved, mothers' warmth was less strong of a predictor for emotional adjustment, and mothers' inductive reasoning also became more contingent on adjustment, suggesting that mothers in such families played more similar roles to fathers.

We also found substantial evidence that both parents showed more desirable patterns of qualitative parenting behaviors - behaviors that are positively associated with children's adjustment - when total joint involvement was higher. When mothers were more involved than fathers, however, mothers' expressed more warmth, whereas fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth). Children's emotional adjustment, too, was lower when mothers were more involved than fathers. These findings highlight the important of considering interactions between quantitative and qualitative involvement, and suggest that all family members might benefit from a more equal distribution of childcare tasks.

With regard to mothers' versus fathers' unique roles in children's emotional development, we found that, in line with the notion of the mother as a warm, supportive attachment figure, maternal warmth and consistency were associated with children's emotional adjustment. Evidence for fathers' unique role was relatively limited; significant within-family correlations between paternal inductive reasoning and emotional adjustment suggest that fathers might play a role in the capacity of disciplinarian. The aforementioned interaction effects between qualitative and quantitative involvement, however, seem to suggest that - above all - the role of fathers is more context dependent than that of mothers.

<!--We used `r cite_r("r-references.bib")` for all our analyses.-->

\newpage

# References
```{r create_r-references}
#r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup