### 2020-04-10 Paper_shortened_pdf.Rmd
Final changes for resubmission

---
title             : "Mothers' and fathers' quantitative and qualitative parenting in relation to children's emotional adjustment: A between- and within-family investigation"
shorttitle        : "INVOLVEMENT AND EMOTIONAL ADJUSTMENT"

author: 
  - name          : "Caspar J. van Lissa"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email         : "c.j.vanlissa@uu.nl"
  - name          : "Renske Keizer"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id            : "2"
    institution   : "Erasmus University Rotterdam Erasmus School of Social and Behavioral Sciences, department of Public Administration and Sociology"

abstract: |
  This longitudinal study of Australian families (n = 1884, from age 6-12)
  examined how fathers' and mothers' quantitative involvement (time spent on
  childcare) and qualitative involvement (specific parenting behaviors) relate
  to children's emotional adjustment development. We used dynamic structural
  equation modeling to disentangle stable between-family differences from
  within-family fluctuations in qualitative parenting and emotional adjustment,
  allowing us to investigate the direction of effects between parents and
  children. Because fathers have been theorized to contribute uniquely to
  emotional adjustment development, we examined differences between mothers and
  fathers, and the interplay between parent- and child sex. We further examined
  whether between-family differences in quantitative involvement,
  operationalized as joint (total) and relative (one parent does more)
  involvement, predict qualitative parenting behaviors and emotional
  adjustment, and moderate the within-family interplay between them. Our
  results primarily revealed between-family correlations of qualitative
  involvement with emotional adjustment. Evidence for within-family effects was
  limited to a parenting effect of maternal warmth, and several child effects,
  which indicated that emotional adjustment evokes changes in parenting.
  Parenting effects varied substantially across families however, especially
  for fathers. Between-family differences in quantitative involvement explained
  some of these differences. When joint quantitative involvement was higher,
  both parents engaged in more desirable qualitative parenting. When mothers
  were relatively more involved than fathers, fathers displayed less desirable
  qualitative parenting, and children were less well-adjusted. There was some
  limited evidence that within-family predictive effects were moderated by
  quantitative involvement, particularly relative involvement. These findings
  suggest that all family members benefit from a more equal childcare
  distribution.

keywords          : "involvement; emotional adjustment; fathers; parenting; longitudinal"
wordcount         : "8634"

bibliography      : ["DSEM_2017.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : yes

class             : "man"
classoption       : "noextraspace"
output            : papaja::apa6_pdf
header-includes   :
    - \raggedbottom
---

```{r load_packages, include = FALSE}
library("papaja")
library(motley)
source("rfunctions.R")
load("results.RData")
load("table_coefficients.RData")
mask <- FALSE
masked <- function(text){
  ifelse(mask, "[masked]", text)
}
```

```{r analysis_preferences}
# Seed for random number generation
set.seed(42)
```

\setlength{\parskip}{0em}

\vspace{\baselineskip}
Children's emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001]. It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013]. Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for their child), and in terms of qualitative parenting behaviors (how parents behave towards their child). Nevertheless, the existing literature has two major shortcomings. Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014]. Yet it is plausible that quantitative involvement might moderate predictive effects between qualitative parenting and emotional adjustment. For example, children's emotional adjustment might be differently impacted by anger coming from a highly involved parent, than from a generally absent parent. Conversely, children's poor emotional adjustment might elicit warm responses from hands-on parents, who are aware of the reasons for their children's emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be disciplined with anger or inductive reasoning. The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.

Secondly, although emotional adjustment is one particular area of development where fathers have been theorized to play a unique role [@paquetteTheorizingFatherchildRelationship2004], fathers are still not consistently included in research [@faganShouldResearchersConceptualize2014]. Consequently, prior research has not sufficiently addressed mothers' and fathers' joint and unique contributions to emotional adjustment development. The present study overcomes this limitation by incorporating both parents. We use sophisticated methods to examine how joint and relative quantitative involvement at age 6 moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment, from the age of 6 to 12 years. We thus obtain a more comprehensive and nuanced understanding of how fathers' and mothers' involvement relate to children's emotional adjustment development.

## Qualitative parenting and emotional adjustment

Early developmental research focused on parental socialization; the notion that parents shape child development through proximal mechanisms, such as qualitative parenting behaviors [@maccobySocializationContextFamily1983; see also @vanlissaRoleFathersMothers2019]. Qualitative parenting behaviors are related to the dimensions underlying parenting styles: Responsiveness and demandingness [@maccobySocializationContextFamily1983]. Substantial evidence indicates that responsiveness, which we operationalized in terms of warm and supportive parenting, is positively associated with children's emotional adjustment [@morrisRoleFamilyContext2007]. Therefore we hypothesized that parental warmth would be positively associated with children's emotional adjustment, $H_{1a}$.<!--@khalequePerceivedParentalAcceptanceRejection2002; -->

The literature linking demandingness and emotional adjustment is less conclusive [@morrisRoleFamilyContext2007]. Demandingness encompasses parenting behaviors related to setting and enforcing expectations for children. Some of these behaviors help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013]. We examined two aspects of demandingness: Consistency; the extent to which consequences for transgressions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments. Consistency correlates positively with emotional adjustment [@brandPerceivedParentalInconsistency1990]. Therefore we also hypothesized that consistency would be positively associated with emotional adjustment, $H_{1b}$. There is less empirical evidence for links between inductive reasoning and emotional adjustment. Given that inductive reasoning promotes the internalization of parental expectations, including display rules for emotions [@grusecImpactParentalDiscipline1994], we hypothesized that it would also be positively related to emotional adjustment, $H_{1c}$.

Negative parenting, which we operationalized as parental anger, is a qualitative parenting dimension not originally subsumed in the parenting styles typology. It is negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999]. Parental anger negatively impacts children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because it is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003]. We thus hypothesized that anger would be negatively related to emotional adjustment, $H_{1d}$.

### Parent or child effects?

The family is commonly viewed as a dynamic system, whose members reciprocally influence one another [@kerrPerceivedParentingStyle2012]. Nevertheless, bi-directional effects are not routinely investigated [@flouriFathersBehaviorsChildren2010]. Most of the research cited in support of hypotheses $H_{1a-d}$ is cross-sectional, and could also support child effects. We thus hypothesize that emotional adjustment positively predicts warmth, consistency, and inductive reasoning, and negatively predicts anger, $H_{2a-d}$. These predictions are in line with studies that investigated child effects specifically, and found that maladjustment elicits less warmth and more negative parenting [@gadeyneLongitudinalRelationsParenting2004]. There is also evidence for bi-directional positive links between parental consistency and emotional adjustment [@lenguaBidirectionalAssociationsTemperament2005]. The adolescent literature provides additional evidence that emotional adjustment positively predicts parental support [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014], and negatively predicts negative parenting [@wijsbroekDirectionEffectsPerceived2011]. Thus, in a way, parents respond positively to emotional adjustment, and negatively to maladjustment.
<!-- Neppl 2009: Adolescents' harsh behavior towards parents predicts harsh parenting, and positive behavior predicts positive parenting -->

### Analyzing the direction of effects

Methods commonly used to examine direction of effects between parents and children have been criticized for failing to disentangle between-family differences from within-family predictive effects [@hamakerCritiqueCrosslaggedPanel2015]. The recently developed random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018], overcome this limitation by disentangling the two levels.

Several studies have re-examined the direction of effects between parents and children using these methods, but mostly in adolescent samples. One such study found that effects of parenting on adjustment that emerged using classical methods became non-significant when stable between-family differences were taken into account [@keijsersParentalMonitoringAdolescent2016]. Another study similarly found that associations between maternal and paternal qualitative parenting and adolescents' emotional adjustment, were mostly explained by between-family level correlations. The remaining predictive effects consisted primarily of child effects, not parenting effects [@vanlissaRoleFathersMothers2019]. The authors theorized that this finding was reflective of the relative equality of parent-child relationships in adolescence. If this is correct, we would expect to find more evidence of parenting effects in the present, younger, sample. These two studies also illustrate that applying these new methods changes the pattern of results in unexpected ways. Consequently, it is impossible to make specific predictions about the pattern of between-family correlations and within-family predictive effects. For now, we will assume that $H_{1a-d}$ and $H_{2a-d}$ apply at both the between- and within-family level.

### Differences between mothers and fathers

The notion that fathers should share child-rearing responsibilities with mothers has become widely accepted in the Western world [@pleckFatherhoodIdealsUnited1997]. Consequently, researchers have come to study qualitative involvement along the same dimensions for mothers and fathers [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004]. Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers [@flouriFathersBehaviorsChildren2010]. Some research suggests that both parents influence children similarly, but that effects might be stronger for mothers [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014]. Others have argued that mothers and fathers might contribute through somewhat different mechanisms. Attachment relationships are grounded in a balance between warmth and security, and exploration [@grossmannDeveloppementAttachementAdaptation1998]. Some have proposed that warmth is more central to mother-child attachment relationships, whereas father-child relationships can be construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013]. By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].

In line with this theory, empirical research has found that mothers' responses to emotional expression are more warm and supportive than fathers' responses, which are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999]. In adolescence, maternal warmth was found to be a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2019]. Similar differentiated associations were found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003]. Finally, there is initial evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is somewhat incongruous with the warm "attachment" relationship typically expected of mothers. Based on this research, we therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demandingness, operationalized as consistency and inductive reasoning, $H_{3}$.

## Quantitative involvement

### Operationalizing quantitative involvement

Early fatherhood research primarily operationalized involvement in *quantitative* terms: Father absence in divorced families, and childcare sharing in dual-earner families [@lambBiosocialPerspectivePaternal1987]. Quantitative involvement has been researched extensively in terms of the absolute amount of childcare provided [see @pleckPaternalInvolvementRevised2004]. However, some have argued that the *division* of childcare tasks is more relevant for children's adjustment [e.g., @keizerInfluenceFathersMothers2019; @craigHowMothersFathers2011]. Indeed, the equity of the childcare distribution is associated with socio-emotional development [@deutschPaternalParticipationChild2001]. Moreover, an unequal distribution often implies that women perform a disproportionate share of the work [@craigDoesFatherCare2006]. This strains the marital relationship [@shockleyItNotWhat2018], and erodes parenting quality, particularly for parents of difficult children [@sheltonInterparentalConflictNegative2008].

There are also methodological advantages to focusing on the distribution of childcare. Operationalizing quantitative involvement in terms of both parents' absolute involvement incurs a potential confound: Effects that appear to be due to fathers' greater involvement might be due to children receiving *more* total care [see @deutschPaternalParticipationChild2001]. This confound can be mitigated by estimating a latent mean and difference [@mcardleLatentVariableModeling2009]. The latent mean reflects joint involvement; the total (average) amount of care provided to children, and tge latent difference reflects relative involvement; the extent to which the mother is more involved than the father (i.e., the childcare distribution).

<!-- examining the effect of the distribution of childcare, while controlling for total care. Some prior research has attempted to achieve this by assessing the childcare distribution directly, by asking which parent is most responsible for certain tasks [@krollEarlyFatherInvolvement2016; ].

Alternatively, when quantitative involvement has been measured in terms of absolute involvement, one can-->

## Main effects of quantitative involvement

### Effects on qualitative involvement

Time spent with children is associated with the quality of parenting behavior [@sabattiniRelationMothersFathers2004]. Indeed, more involved parents tend to experience greater parenting self-efficacy, possibly reflecting the effects of experience [@gialloMothersFathersInvolvement2013a]. Furthermore, when both parents are more involved, they monitor each other's parenting behaviors to ensure that children are raised well [@mclanahanConsequencesFatherAbsence2014]. Consequently, we hypothesize that greater joint involvement will positively predict both parents' desirable parenting behaviors (warmth, consistency, and inductive reasoning), and negatively predict parental anger, $H_{4a-d}$.

By the same reasoning, we might hypothesize that greater differential involvement would positively predict the desirable parenting behaviors of the more involved parent, and negatively predict those of the less involved parent, $H_{5a-d}$ [@sabattiniRelationMothersFathers2004]. However, there are also reasons to expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment. Retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004]. One reason might be that parents are less able to monitor each other's child-rearing behavior when they are not equally involved [@mclanahanConsequencesFatherAbsence2014]. Additionally, differential involvement might negatively impact the marital relationship, and thereby undermine both parents' involvement quality. Childcare responsibilities are a heavy burden [@deutschPaternalParticipationChild2001]. An unfair distribution strains the interparental relationship, interferes with sensitive parenting, and undermines child well-being [@davies2006interparental]. When care is distributed relatively equally, mothers tend to be more satisfied [@pharesFathersMothersInvolvement2009], and the family system more harmonious. Consequently, balanced co-parenting is associated with greater parental warmth [@barnettInterdependenceParentingMothers2008].

### Effects on emotional adjustment

Bioecological theories argue that children require a certain amount of care; who provides it is somewhat flexible [see @bronfenbrennerEcologyHumanDevelopment1979]. We therefore hypothesize that greater joint care should positively predict children's emotional adjustment, $H_{6a}$. Further, it has been argued that children feel less supported when they perceive that one of their parents is relatively uninvolved, than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001]. We therefore predict that greater relative involvement will negatively predict emotional adjustment, $H_{6b}$.

## Interactions between involvement quality and quantity

Interactions between involvement quality and quantity have largely been overlooked in the parenting literature [@jaffeeLifeFatherBenefits2003]. Consequently, it is not possible to substantiate specific hypotheses for such interactions. One notable exception is Feinberg's coparenting model [@feinbergInternalStructureEcological2003]. Coparenting refers to the ways parents relate to each other in the parenting role. Although the coparenting literature is typically invoked to describe child-rearing in divorced families, it also holds insights for intact families. According to Feinberg, the "division of labor" in childcare, analogous to our operationalization of relative involvement, affects qualitative child-rearing behaviors and children's adjustment directly, and moderates the link between these two constructs. Applying this insight from the coparenting literature, we might expect parents' quantitative involvement to moderate effects between qualitative parenting behaviors and emotional adjustment. Due to the limited literature, it is not possible to formulate specific hypotheses about the nature of these moderating effects. There is, however, preliminary evidence to indicate that interactions between involvement quality and quantity are relevant. For instance, fathers' playful and caring involvement positively predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@jiaTransactionalRelationsFather2012]. Furthermore, the quantity of paternal school-related involvement interacted with paternal warmth to positively predict adolescents' academic achievement and behavioral adjustment [@wangParentalInvolvementAfrican2014]. We thus explored the moderating effects of joint and relative quantitative involvement on the interplay between qualitative parenting behaviors and emotional adjustment.

## Child sex differences

Child sex differences are considered to be an important covariate in the literature on parenting and emotional adjustment. Parents raise boys and girls differently, expressing more demandingness with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2019], and more anger with boys [@strausCorporalPunishmentAmerican1999]. Furthermore, the same-sex parent is thought to play a stronger role in socialization [@busseySocialCognitiveTheory1999]. Finally, mounting evidence suggests that parental warmth is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2019]. We therefore hypothesize that parents will display more warmth, consistency, and inductive reasoning, and less anger, towards girls than boys, $H_{7a-d}$. We also predict associations between qualitative parenting and emotional adjustment to be stronger between children and their same-sex parent, $H_{8a-d}$. Finally, although sex differences in emotional adjustment have been found in older children [@vanlissaRoleFathersMothers2019], there is little evidence for such differences in young children [@else-questGenderDifferencesTemperament2006a].

## The present study

The present study examined whether specific qualitative parenting behaviors are associated with children's emotional adjustment, and whether these associations are moderated by quantitative involvement. We aimed to make several contributions. First, by including both mothers and fathers, we can tease apart their common and unique roles in children's emotional adjustment development. Second, by incorporating both qualitative and quantitative measures of involvement, we can examine interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work. Third, by disentangling stable between-family differences from dynamic within-family predictive processes, we can establish at which level associations occur, and assess the within-family direction of effects between qualitative involvement and emotional adjustment. Finally, we examined child sex as a moderator of between-family differences in all study variables and within-family predictive effects.

# Methods

## Participants

We employed data from the `r masked("K-cohort")` of the `r masked("Longitudinal Study of Australian Children (LSAC)")`, waves `r masked("2-5")`, which were collected by means of biannual home visits, and self-report mail questionnaires, from 2006-2012. Each wave was reviewed and approved by the
`r masked("Australian Institute of Family Studies Ethics Committee")`, which is a Human Research
Ethics Committee registered with the `r masked("National Health and Medical Research Council")`. Documentation of the ethics approval is available online [@australianinstituteoffamilystudiesGrowingAustralia2020]. `r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population. Complete details regarding the study design, sample, and procedure are available online [@australianinstituteoffamilystudiesGrowingAustralia2020].

From the total sample (4.983 focal children at the first wave), we selected the 1884 native Australian, non-minority, intact biological families that constituted our target population. Although minority- or immigration status and family composition are important factors, they are beyond the scope of this paper. At the first wave, children in our sample were approximately 6 years old, $M = 6.28$, $SD = .45$, and 12 at the final wave ($M = 12.42$, $SD = .49$); on average, mothers were 34.89 ($SD = 4.71$), and fathers 36.94 ($SD = 5.48$). Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$); the remainder were single-earner families($N = 669$ father-only, $N = 41$ mother-only). The LSAC did not measure income as a continuous variable, but as an ordered categorical variable. For fathers, the modal category was "\$26,000-\$51,999" ($N = 663$ fathers reported this value), and for mothers, "\$25,999 or less per year" ($N = 1152$ reported this value).

## Measures

We calculated mean scores for all study variables with multiple indicators. Scale descriptive statistics, number of items, attrition (which ranged from 7-14% across respondents), reliability, and intraclass correlations (ICCs) are reported in Supplementary Table 1; scale correlations are provided in Supplementary Table 2. The difference in fit between structurally invariant models and metrically invariant models was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3). We thus deemed it defensible to assume metric measurement invariance (i.e., to assume that the factor loadings of each construct are the same across time points and respondents).

### Qualitative involvement.

Responsiveness and inductive reasoning were measured using the warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire [CRQ, @patersonAssociationBehaviouralAdjustment2001]. Consistency and negative parenting were measured using the consistency (e.g., "How often is this child able to get out of punishment when he/she really sets his/her mind to it?", reverse coded) and anger (e.g., "How often are you angry when you punish this child?") scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000]. These self-report instruments used five-point scales, ranging from "never or almost never" to "always or almost always".

### Quantitative involvement.

Measures of quantitative involvement varied from wave to wave, due to children's changing care needs [@wangATUSFathersInvolvement2009]. We used the first wave reports as an indicator of between-family differences in quantitative involvement. Both parents rated their involvement in four different child care activities on a five-point scale ranging from "once a day or more" to "not at all" [e.g., "In the past month how often did you help this child brush his/her teeth?"; @baxterFatheringAustraliaCouple2011].

### Emotional adjustment.

Emotional adjustment was measured using the reverse-coded emotional distress subscale of the teacher-reported Strengths and Difficulties Questionnaire, on a three-point scale from "not true" to "certainly true" [e.g., "Often unhappy, downhearted or tearful"; @goodmanPsychometricPropertiesStrengths2001].

## Strategy of analyses

To address our research questions, we used multilevel, autoregressive structural equation modeling [@asparouhovDynamicStructuralEquation2018]. In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean). The ICCs (supplementary Table S1) indicate the proportion of variance at the between- versus within-family level. For warmth, consistency, and anger, the ICCs were high, ranging from [.60, .64]. This indicates that about 60% of the variance in these parenting behaviors is due to between-family differences. For inductive reasoning, they were notably lower, [.46, .50]. For emotional adjustment, they were even lower, $r = .25$.

Disentangling between- and within-family level variance allows us to determine at which of these levels associations occur, and to estimate the direction of within-family effects more accurately [@hamakerCritiqueCrosslaggedPanel2015]. Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between random intercepts. Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects. In this way, we included joint and relative quantitative involvement, and child sex as moderators. Figure \@ref(fig:figure1) shows a graphical representation of the resulting model. We conducted separate analyses for the four qualitative parenting dimensions, because a model including all four variables at once was too complex to estimate.

```{r figure1, out.width="100%", fig.cap="T-1: previous measurement occasion. ri: random intercept. s: random slope. M: Mother's qualitative involvement. E: Emotional adjustment. F: Father's qualitative involvement. inv: quantitative involvement. Thus, sME represents the random slope of Mothers' qualitative involvement on emotional adjustment."}
knitr::include_graphics("Figure1.png")
```

Analyses were conducted using Mplus 8.0 and the R-package MplusAutomation. This type of multilevel autoregressive model can only be estimated using Bayesian methods. We used two Gibbs-sampler chains of 50000 iterations each. Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing-at-random (a non-parametric MCAR test resulted in $p > .05$). We used mildly informative inverse Wishart priors for covariance matrices, and normal priors centered on zero for regression coefficients. We assessed convergence through visual inspection of the chains and posterior distributions. We report the posterior median, and a 95% highest posterior density interval (HPDI). This is the "narrowest" possible interval that contains 95% of samples from the posterior distribution. It represents the most likely population values of the parameter in question. Results are best interpreted by examining the effect size (correlations and standardized regression coefficients), and the width of the HPDI. Significance is not a Bayesian concept, but we use the shorthand "significant" to refer to estimates for which less than 5% of the posterior distribution had the opposite sign from the estimate. Supplementary Table S4 reports all standardized and unstandardized model parameters. All code and output are available online at [masked].

# Results

## Is qualitative parenting correlated with emotional adjustment?

We hypothesized that warmth, consistency, and inductive reasoning would correlate positively, and anger negatively, with emotional adjustment. At the between-family level, we examined correlations between families' average levels of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabcors)). These correlations indicate whether between-family differences in the average levels of these parenting behaviors are associated with between-family differences in children's emotional adjustment. Counter to hypotheses $H_{1,2a}$, we found no significant correlations between parental warmth and children's emotional adjustment. In support of hypotheses $H_{1,2b}$, we found significant positive correlations between maternal and paternal consistency, and emotional adjustment. Contrary to hypotheses $H_{1,2c}$, we found that maternal and paternal inductive reasoning were negatively correlated with emotional adjustment. Finally, as hypothesized in $H_{1,2d}$, we found significant negative correlations between maternal and paternal anger, and emotional adjustment. To sum up, in families where parents reported more consistency, children displayed better emotional adjustment - and in families where parents reported more inductive reasoning or anger, children showed poorer emotional adjustment.

```{r tabcors, results = "asis"}
cor_tables <- lapply(results, corTable)

table_between_cor <- lapply(cor_tables, function(x){x[["Between"]][3:5, 3:5]})
table_between_cor <- order_by_name(table_between_cor)
table_between_cor <- do.call(cbind, table_between_cor)
table_between_cor <- cbind(rename_pars(rownames(table_between_cor)), table_between_cor)
table_between_cor <- as.matrix(table_between_cor)
colnames(table_between_cor) <- c("Parameter", rep(table_between_cor[,1], 4))
rownames(table_between_cor) <- NULL
table_between_cor <- gsub("\\*+", "*", table_between_cor)
table_between_cor <- gsub("^1$", "", table_between_cor)
table_between_cor <- table_between_cor[-3, -seq(2, 11, by = 3)]
table_between_cor <- rbind(c("Between-family level correlations", rep(NA, ncol(table_between_cor)-1)), table_between_cor)

table_within_cor <- lapply(cor_tables, `[[`, "Within.Std.Averaged.Over.Clusters")
table_within_cor <- order_by_name(table_within_cor)
table_within_cor <- do.call(cbind, table_within_cor)
table_within_cor <- cbind(rename_pars(rownames(table_within_cor)), table_within_cor)
table_within_cor <- as.matrix(table_within_cor)
colnames(table_within_cor) <- c("Parameter", rep(table_within_cor[,1], 4))
rownames(table_within_cor) <- NULL
table_within_cor <- gsub("\\*+", "*", table_within_cor)
table_within_cor <- gsub("^1$", "", table_within_cor)
table_within_cor <- table_within_cor[-3, -seq(2, 11, by = 3)]
table_within_cor <- rbind(c("Within-families level correlations", rep(NA, ncol(table_within_cor)-1)), table_within_cor)

table_cors <- rbind(table_between_cor, table_within_cor)
write.csv(table_cors, "table_cors.csv", row.names = FALSE)
#table_cors[is.na(table_cors)] <- ""
rownames(table_cors) <- NULL
colnames(table_cors)[1] <- "Param."
#reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Param.: Parameter. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

We also examined within-family correlations, which are based on within-family fluctuations in the study variables (Table \@ref(tab:tabcors)). These correlations indicate whether fluctuations in parents' parenting behaviors are correlated with fluctuations in their children's emotional adjustment. In support of hypothesis $H_{1a}$, we found significant positive within-family level associations between fathers' and mothers' warmth, and emotional adjustment. This means that if, in any particular wave, parents report more warmth than usual, their children displayed better emotional adjustment than usual. Furthermore, in partial support of hypotheses $H_{1b,c}$, we found significant positive correlations of emotional adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning. Finally, contrary to hypothesis $H_{1d}$, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger. Counter to hypothesis $H_{3}$, none of the differences between mothers and fathers in the strength of either between- or within-family level correlations were significant (see Supplementary Table S4).

## Do within-family fluctuations in parenting predict emotional adjustment or vice versa?

In addition to the aforementioned within-family correlations, we also estimated over-time predictive effects between within-family fluctuations in qualitative parenting, and emotional adjustment. Each family has family-specific regression coefficients, and across families, we estimate the intercept and variance of these cross-lagged effects. The intercept indicates the average of the within-family predictive effects across families, and the variance indicate whether the family-specific regression coefficient varies in strength (and sign) across families. Hypotheses $H_{1a-d}$ also apply to these within-family predictive effects. To test these hypotheses, we examine the average standardized regression coefficients (see Table \@ref(tab:tabcrosspath)). With regard to parenting effects, we found that, in line with hypothesis $H_{1a}$, maternal warmth positively predicted children's emotional adjustment in the next wave. The standardized regression coefficient of this effect is interpreted as follows: On average across families, a one standard deviation increase in maternal warmth in one wave predicts a $.99$ standard deviation increase in children's emotional adjustment in the next wave. We found no evidence for hypotheses $H_{1b-d}$ in terms of parenting effects. The variances of all parenting effects were significant, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families. Moreover, these variances were all significantly larger for fathers than for mothers (see Supplementary Table S4), indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.

```{r tabcrosspath, results = "asis"}
tabparenteff <- rbind(tab_std_sections$`Intercepts of parenting effects`, 
                      tab_unst_sections$`Between-family level residual variances`[c(1, 5:6),])

tabparenteff$Parameter <- gsub("^\\(.+on (.+)\\)$", "\\1", tabparenteff$Parameter)
tabparenteff$Parameter[c(1,4)] <- c("Average parenting effect of:", "Variance of parenting effect of:")

tabchildeff <- rbind(tab_std_sections$`Intercepts of child effects`,
                      tab_unst_sections$`Between-family level residual variances`[c(1, 7:8),])

tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
tabchildeff$Parameter[c(1,4)] <- c("Average child effect on:", "Variance of child effect on:")
tab <- rbind(tabparenteff, tabchildeff)
#tab[is.na(tab)] <- ""
rownames(tab) <- NULL

# Add letters for contrast tests
letter <- 1
tab <- add_letters(list("Mother effect - child effect on mother" = c(2, 8),
                 "Father effect - child effect on father" = c(3, 9)),
            tab, tab_unst_sections)

tab <- add_letters(list("F-M Parent effects on emo." = c(2, 3),
                        "F-M Variance of parenting effect" = c(5,6),
                 "F-M Child effects on parenting" = c(8, 9),
                 "F-M Variance of child effect" = c(11, 12)),
            tab, tab_unst_sections)

tab$Parameter <- gsub("^Variance", "Unstandardized variance", tab$Parameter)
names(tab)[1] <- "Param."
reg_table(tab, caption = "Standardized between-family level intercepts (average effect) and unstandardized variances of within-family predictive effects.", note = "Param.: Parameter. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate. Matching letters indicate that parameters differed significantly; see Supplementary Table S4 for all difference tests.")
```

With regard to child effects, in line with hypothesis $H_{2a}$, the average cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabcrosspath)). Thus, on average, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more warmth than usual in the next wave. Contrary to hypothesis $H_{2b}$, the average effect of emotional adjustment on maternal consistency was non-significant. As hypothesized in $H_{2c}$, the intercept for maternal inductive reasoning was positive. Counter to hypothesis $H_{2d}$, the intercept for maternal anger was positive, not negative. Thus, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more anger than usual in the next wave. In support of hypothesis $H_{2a}$, the average child effect of adjustment on paternal warmth was positive and significant. None of the other intercepts ($H_{2b-d}$) for child effects on fathers' parenting were significant, suggesting that fathers' parenting was relatively unresponsive to fluctuations in children's emotional adjustment. The variances of all child effects were significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs somewhat between families.

We established direction of effects by testing the difference between corresponding parent- and child effects (see Table \@ref(tab:tabcrosspath) and Supplementary Table S4). All effects that were significant in one direction, and non-significant in the opposite direction, had significant difference tests. All effects that were significant (or non-significant) in both directions had non-significant difference tests. We examined differences between fathers and mothers in parent- and child effects the same way. All effects that were significant for one parent but not significant for the other parent had significant difference tests, except for the child effect of emotional adjustment on parental anger, which did not differ significantly between parents.

## Does quantitative involvement predict the average level of parenting and adjustment?

To investigate our hypotheses $H_{4-6}$, we examined the predictive effects of joint and relative quantitative involvement on the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). With regard to effects of joint quantitative involvement on parenting, we found partial support for our hypotheses $H_{4a-c}$: For both parents, greater joint involvement predicted increased warmth, consistency, and inductive reasoning. Notably, the association between joint involvement and warmth was almost double in size for fathers, as compared to mothers. We did not find the hypothesized negative effects of joint involvement on parental anger, $H_{4d}$. With regard to effects of relative quantitative involvement on parenting, we found limited support for hypotheses $H_{5a-d}$, that greater differential involvement would positively predict the desirable parenting behaviors of the more involved parent, and negatively predict those of the less involved parent. For mothers, the only significant association we found was that mothers displayed more warmth when they were relatively more involved than fathers. We found that fathers displayed less warmth, consistency, and inductive reasoning, when mothers were relatively more involved than fathers. With regard to effects of quantitative involvement on emotional adjustment, counter to hypothesis $H_{6a}$, we found that greater joint involvement negatively predicted emotional adjustment. Thus, children of highly involved parents generally displayed poorer emotional adjustment. Hypothesis $H_{6b}$ was supported: Children's emotional adjustment was lower in families where mothers were relatively more involved than fathers.

```{r tabinvolvement, results = "asis"}
tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
tabinv$Parameter[c(1,5)] <- c("Effects of joint quantitative involvement on random intercepts of:",
                              "Effects of relative quantitative involvement on random intercepts of:")
tabinv$Parameter <- gsub("^(.+) on (?!random).*$", "\\1", tabinv$Parameter, perl= TRUE)
#reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
tabinvmod <- rbind(
  tab_std_sections$`Moderating effect of quantitative involvement on parenting effects`[c(1:3, 1, 4:5),],
  tab_std_sections$`Moderating effect of quantitative involvement on child effects`[c(1:3, 1, 4:5),])
tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on parenting effects of:",
                              "Moderating effects of relative involvement on parenting effects of:",
                              "Moderating effects of joint involvement on child effects on:",
                              "Moderating effects of relative involvement on child effects on:")
tab <- rbind(tabinv, tabinvmod)
#tab[is.na(tab)] <- ""
rownames(tab) <- NULL
names(tab)[1] <- "Param."
cat(as.character(reg_table(tab, caption = "Standardized regression coefficients of quantitative involvement on between-family differences in qualitative parenting and emotional adjustment, and on within-family predictive effects", note = "The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1. Param.: Parameter. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")))
```

## Does quantitative involvement moderate within-family predictive effects?

We conducted exploratory analyses of interactions between involvement quality and quantity, by investigating whether between-family differences in quantitative involvement had significant effects on the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). The novelty of this research area precluded us from formulating specific hypotheses. With regard to parenting effects, the only significant moderation we found was for maternal warmth: The positive effect of maternal warmth on adjustment was less pronounced in families in which both parents were highly involved. Specifically, a $+1SD$ increase in joint involvement predicted a $-.14SD$ decrease in the within-families cross-lagged effect of maternal warmth on emotional adjustment (see Table \@ref(tab:tabcrosspath)). This same within-families effect increased by $.11$ of a standard deviation of the effect when relative involvement was one standard deviation higher, i.e., when mothers were more involved than fathers. We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.

With regard to the moderation of child effects by quantitative involvement, we found that greater joint quantitative involvement, and mothers' greater relative involvement, both negatively moderated the within-family effect of emotional adjustment on maternal warmth. We found no effects for consistency. For inductive reasoning, we found that greater joint involvement negatively moderated the within-family effect of adjustment on maternal inductive reasoning. Greater relative involvement also negatively moderated the effect of emotional adjustment on fathers' inductive reasoning. Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment predicted a stronger decrease in paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning. We found no moderating effects for anger.

## Child sex differences

We found sex differences in the average levels of qualitative parenting behaviors: In partial support of $H_{7a,d}$, fathers displayed more warmth towards daughters, and both parents displayed more anger towards sons (Table \@ref(tab:tabsex)). Counter to hypotheses $H_{7b,c}$, there were no sex differences in consistency, and both parents displayed more inductive reasoning towards sons. We found no sex differences in the average level of emotional adjustment. Hypotheses $H_{8a-d}$ were rejected: There was no evidence for stronger associations between qualitative parenting and emotional adjustment between children and their same-sex parent. The only predictive effects moderated by child sex were child effects of emotional adjustment on maternal and paternal inductive reasoning: When children displayed worse emotional adjustment than usual in one wave, their parents reported stronger increases in inductive reasoning in the next wave for boys than for girls.

```{r tabsex, results = "asis"}
tabsex <- rbind(
  tab_std_sections$`Sex differences in parenting and adjustment`,
  tab_std_sections$`Sex differences in cross-lagged effects`[c(1:3, 1, 4:5),])
tabsex$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabsex$Parameter)
tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
tabsex$Parameter[c(1,5,8)] <- c("Sex differences in random intercepts of parenting and adjustment",
                              "Moderation of parenting effects by sex",
                              "Moderation of child effects by sex")
#tabsex[is.na(tabsex)] <- ""
rownames(tabsex) <- NULL
names(tab)[1] <- "Param."
reg_table(tabsex, caption = "Standardized regression coefficients of child sex on average levels of qualitative parenting and emotional adjustment, and on within-family predictive effects.", note = "Child sex is a binary predictor (-1 = female, 1 = male). The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1. Param.: Parameter. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

# Discussion
<!-- To recapitulate, we set out to investigate how fathers' and mothers' quantitative and qualitative involvement relate to children's emotional adjustment development. We used innovative methods to disentangle stable between-family differences from within-family fluctuations. 
-->
To recapitulate, we set out to obtain a more comprehensive and nuanced understanding of how fathers' and mothers' quantitative and qualitative involvement relate to children's emotional adjustment development. We used innovative methods to disentangle stable between-family differences from within-family fluctuations in the study variables, allowing us to investigate at which of these level associations occur, and establish the direction of effects between parents and children. Furthermore, we examined how joint and relative quantitative involvement predict parenting and emotional adjustment, and moderate the longitudinal interplay between qualitative parenting and emotional adjustment.

## Correlations between qualitative parenting and emotional adjustment

We found mixed evidence for our hypotheses that warmth, consistency, and inductive reasoning would correlate positively, and anger negatively, with emotional adjustment. Parental warmth correlated with emotional adjustment only at the within-family level. Consistency correlated positively with adjustment at the between-family level for both parents, and at the within-family level for mothers. Correlations with inductive reasoning were, paradoxically, opposite in sign at the between- and within-family levels. At the between-family level, parents who engaged in more inductive reasoning had children with poorer emotional adjustment. At the within-family level, fathers' inductive reasoning was positively correlated with adjustment. Finally, for anger, we found the predicted negative associations at the between-family level only, which means that, in families where mothers or fathers display high anger, children tend to be less emotionally well-adjusted. Most of these findings are in line with hypotheses based on prior research that did not disentangle between- and within-family variance. However, the few notable deviations from the expected pattern illustrate an important insight: Correlations do not necessarily occur at both of these levels, and can even be opposite in sign; a phenomenon known as the ecological fallacy [see @keijsersParentalMonitoringAdolescent2016].

A potential explanation for the paradoxical findings for inductive reasoning might be that the construct of inductive reasoning confounds *how* parents discipline their children (sample item: "How often do you talk it over and reason with this child..."), with *how often* their child misbehaves ("...when he/she misbehaved?"). Stable between-family differences in inductive reasoning might tap into *how often* the child misbehaves. The negative between-family correlations between inductive reasoning and emotional adjustment suggest that poorly adjusted children are disciplined more often. Within-family fluctuations, by contrast, might tap into context-specific changes in *how* parents discipline their children to a greater extent [@critchleyParentDisciplinaryBehavior2006]. When parents use more inductive reasoning than they usually do, talking over and reasoning about misbehavior, this might benefit emotional adjustment.

## Direction of effects

In terms of the direction of within-family predictive effects, we found within-family predictive effects between parental warmth and emotional adjustment, which were bi-directional for mothers. This suggests that parental warmth and children's emotional adjustment reciprocally influence one another within families; potentially through causal mechanisms. There was no other evidence for parenting effects however. Child effects outweighed parenting effects both in number and in size. The direction of effects thus appears to flow primarily from children to parents.

With regard to these child effects, on average, children who displayed better emotional adjustment in one wave evoked more warmth from both parents, and more inductive reasoning from mothers in the next wave. This latter finding is in line with the aforementioned within-family correlations for inductive reasoning. Counter to our hypothesis, we found that children who displayed better emotional adjustment in one wave evoked greater maternal anger in the next wave. One possible explanation is regression to the mean: On average, emotional adjustment showed no-to-negative stability coefficients in all models. That means that better emotional adjustment in one wave, was likely to be followed worse adjustment in the next wave. Looking beyond direction of effects, it is worth noting that the brunt of negative associations between anger and adjustment appears to be due to stable between-family differences. The sparsity of within-family effects of anger might be due to a ceiling effect: If the stable level of parental anger in a family is high enough to negatively impact children's emotional adjustment, over-time fluctuations of that anger might no longer make much of a difference.

## Differences between mothers and fathers

Correlations of parenting behaviors with children's adjustment were all similar for mothers and fathers at the between-family level. Within-family correlations, however, were stronger for mothers' consistency, and stronger for fathers' warmth and reasoning. As mentioned before, mothers tend to spend more time with their children than fathers; also in our sample. This might explain why associations for maternal consistency were stronger: Spending more time with the child provides greater opportunity for consistent behavior. For fathers, associations were stronger for inductive reasoning and warmth - behaviors which can also be expressed in one-off interactions, such as when  disciplining a child (inductive reasoning), or during play time (warmth). The stronger association with inductive reasoning for fathers is in line with the classic view of the father as disciplinarian [@flouriFathersBehaviorsChildren2010].

With regard to average parenting effects, the only difference between mothers and fathers was the predictive effect of warmth on emotional adjustment, which was evident only for mothers. This is in line with prior work, indicating that warmth is an important socialization mechanism for mothers [@vanlissaRoleFathersMothers2019]. All parenting effects varied significantly across families, however,  and this variability was notably greater for fathers than mothers. This suggests that fathers' influence is more context-dependent than mothers' influence, and highlights the importance of exploring predictors of between-family differences in within-family processes in future research.

With regard to child effects, we found substantial evidence that mothers' parenting was more responsive to children's emotional adjustment than fathers' parenting. Specifically, mothers' anger, inductive reasoning, and warmth were predicted by adjustment, whereas these effects were absent for fathers. This is in line with some prior studies, which found that mothers are more sensitive to children's needs [see @faganShouldResearchersConceptualize2014, p. 395].

## Does quantitative involvement predict qualitative parenting behaviors?

Turning to our findings from linkages between quantitative involvement and qualitative parenting behaviors, our results were in line with our hypothesis that greater involvement would be associated with more desirable parenting behaviors [@sabattiniRelationMothersFathers2004]. Our results revealed that, when both parents were more involved, they displayed greater warmth, consistency, and inductive reasoning. Fathers, in particular, reported substantially greater parental warmth when both parents were more involved. We also found that, when mothers performed more care tasks than fathers, mothers displayed more warmth, and fathers displayed less warmth, consistency, and inductive reasoning. This increase in maternal warmth might reflect an attempt to compensate for fathers' absence [@shinnFatherAbsenceChildren1978]. These results indicate that, when childcare is unequally distributed, fathers engage in less desirable qualitative parenting behaviors [see also @craigHowMothersFathers2011].

## Quantitative involvement and emotional adjustment

Our operationalization of quantitative involvement, in terms of joint involvement and relative involvement, allowed us to addresses the claim that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001]. However, counter to our hypothesis, joint involvement was *negatively* associated with between-family differences in emotional adjustment. One potential explanation is suggested by self-determination theory: Excessive levels of involvement might undermine children's autonomy needs, thereby leading to increased emotional maladjustment [@vandergiessenPerceivedAutonomySupport2014]. Research on "helicopter parenting" supports this interpretation, as psychological well-being was found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014]. Another potential explanation is that the amount of quantitative care provided is driven, in part, by children's objective needs, such as atypical or delayed development. Such factors require increased quantitative care, and are also associated with children's emotional maladjustment, which could result in a negative correlation between the two.

In line with predictions derived from coparenting theory, we found that children's emotional adjustment was lower in families where mothers were relatively more involved than fathers - even after controlling for the joint amount of care provided. Although substantial empirical work has indicated that the marital relationship suffers when childcare is unequally distributed [@shockleyItNotWhat2018], our research is one of few to examine the implications for children's adjustment [see also @deutschPaternalParticipationChild2001]. We caution against undue causal interpretation of these findings, however. Future research is required to examine potential pathways, such as whether children's adjustment suffers when they feel like both parents are not equally invested in their well-being, or whether fathers are more likely to withdraw from childcare responsibilities when children display poor emotional adjustment, or whether differential involvement strains the marital relationship, and this strain spills over to children's emotional adjustment.

## Does quantitative involvement moderate links between qualitative parenting and emotional adjustment?

We found limited evidence that quantitative involvement moderates parenting effects on emotional adjustment. Specifically, joint- and relative quantitative involvement moderated the effect of maternal warmth on adjustment: Mothers' warmth predicted emotional adjustment less strongly when both parents were more involved, and more strongly when mothers were relatively more involved than fathers. These findings suggests a compensatory effect, whereby mothers' warmth can make up for disadvantages incurred by having a relatively uninvolved father [@shinnFatherAbsenceChildren1978].

Evidence that quantitative involvement moderates child effects was somewhat stronger. Firstly, when joint involvement was higher, and when mothers' relative involvement was greater compared to fathers, emotional adjustment became less predictive of maternal warmth. These findings can be interpreted in the context of the sizable average positive effect of adjustment on maternal warmth. This positive effect was somewhat smaller in families where mothers' involvement was high - either in terms of joint, or relative involvement. Secondly, joint quantitative involvement positively moderated the link from emotional adjustment to maternal inductive reasoning. This suggests that mothers have more of a role as disciplinarian in families where both parents are highly involved. When mothers were relatively more involved than fathers, fathers decreased their inductive reasoning more strongly in response to children's positive adjustment, and increased it more strongly in response to maladjustment. Together, these findings suggests that fathers take more of a role as disciplinarian when mothers are relatively more involved in care tasks.

## Child sex differences

In line with prior research, we found indications that parents are stricter with sons than with daughters; specifically, parents displayed more anger and inductive reasoning, and less warmth, towards sons. Sex also moderated child effects: Emotional adjustment predicted both parents' inductive reasoning more negatively when children were male. This finding suggests that some of parents' greater strictness toward boys is alleviated when boys are well-adjusted.

## Theoretical implications of disentangling within- and between-family variance

The present study mostly found evidence for correlations at the between-family level; evidence for within-family predictive effects was limited, and revealed mostly child effects. In a previous study with a comparable design, conducted in an adolescent sample, we similarly found mostly between-family level correlations and child effects [@vanlissaRoleFathersMothers2019; see also @keijsersParentalMonitoringAdolescent2016]. We explained that finding in light of the relative equality of parent-adolescent relationships. We theorized that the between-family level correlations might reflect the consolidated influence of within-family parent effects that took place at an earlier age, and hypothesized that parenting effects would be more prevalent, and between-family correlations weaker, in younger children. The present findings refute these conjectures. The preponderance of child effects and between-family level correlations in a sample of younger children, age 6-12, suggests that either parent effects occurred at an even younger age, or that something else is at play.

As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory. We should reconsider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it? Our findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time. Our findings further serve as a reminder that we should be critical of causal interpretations of research that does not account for stable between-family differences [e.g., @brandPerceivedParentalInconsistency1990].

Future research should also devote greater attention to between-family level correlations as an outcome of interest. The intraclass correlations of parenting behaviors were high in our study, meaning that the brunt of variance in parenting behaviors is due to stable between-family differences. Between-family level correlations can reflect consolidated effects of parenting effects at an early age [@vanlissaRoleFathersMothers2019], or alternatively, they can reflect the influence of third variables that impact both constructs. We should seek to understand why between-family level correlations are so prominent by comparing between-family correlations across children of different ages, including very young children, and by examining the influence of between-family covariates, including sociological factors. For example, economic hardship undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990]. Due to disciplinary differences, such sociological factors are rarely examined in conjunction with sophisticated models of reciprocal influences between parents and children [see also: @keizerRoleFathersIntergenerational2020].

Methodological innovations that examine the role of stable between-family differences can thus inspire more sophisticated theoretical conceptualizations of parental influences in child development. At the same time, we should consider the limitations of these methods. The assumption that within-family fluctuations adequately capture all causal influences that parents exert on their children, for instance, is questionable. Stable factors can also exert changing influences over time. As an illustrative example: The genetic makeup that is passed from parents to children is stable, and yet the expression of some genes emerges over time. It might similarly be possible for stable parental psychosocial characteristics to influence children differently at different time points. Another important limitation is that within-family cross-lagged effects are time scale dependent. If parents shape their children through specific behaviors in dyadic interactions, then annual reflective self-reports might fail to capture parenting effects. Thus, although these methodological innovations are undoubtedly a step forward, they might not be the definitive paradigm for investigating the interplay between parents and children.

## Strengths and Limitations

Several strengths allow us to make unique contributions to the literature. First, mothers and fathers both completed identical parenting measures, which allowed us to disentangle and compare their unique roles in children's emotional adjustment development. Second, we incorporated both quantitative and qualitative involvement. This allowed us to provide a more nuanced overview of the links between parental involvement and emotional adjustment development than prior work, including how the distribution of quantitative childcare relates to parenting quality and emotional adjustment, and moderates links between these two constructs. Although some prior studies have also incorporated both quality and quantity of parenting, this work did not disentangle the total amount of care children received from the relative distribution of care tasks [see @deutschPaternalParticipationChild2001]. We addressed this limitation by estimating the latent mean and difference of absolute maternal and paternal quantitative involvement. This allows us to disentangle the effect of joint involvement, which reflects the total amount of care children receive, from relative involvement, which reflects the childcare distribution. Third, we used innovative methods to examine the extent to which parenting and adjustment are linked at the between- versus within-family level. Compared to methods that do not distinguish between these levels, our approach estimates dynamic reciprocal processes within the family more accurately. Our findings contribute to a growing literature suggesting that stable between-family differences might play a larger role than was previously thought, even in early childhood. Confidence in the results is bolstered by the fact that we used multi-informant reports: Each parent reported on their own parenting, and teachers reported on children's emotional adjustment. This avoids common method bias and single reporter bias. As teachers are likely to be "blind" to parents' child-rearing behavior, and have no stake in the parent-child relationship, their reports might be less biased in general.

A potential limitation of the present study is that teachers' reports of emotional adjustment might be more noisy than self- or parent-reports. A teacher is responsible for many pupils, and might not be aware of each pupil's emotional adjustment. Moreover, although the instrument we used assessed *overt* symptoms of emotional maladjustment, people are generally poor judges of others' emotional states [@ickesEmpathicAccuracy1997]. Consequently, our effect sizes might be attenuated by noisy measures of emotional adjustment. Teachers' reports do have clear benefits as well: Young children might lack the ability to report accurately on their own emotional adjustment, and using clinical assessments might be unfeasible for large-scale panel studies. Another limitation to consider is the fact that we operationalized qualitative parenting in terms of specific parenting behaviors, whereas others have focused on parenting *styles* composed of specific combinations of these behaviors [@maccobySocializationContextFamily1983]. We focused on specific parenting behaviors for two reasons. First, although aggregating specific parenting behaviors into parenting styles is theoretically defensible, it also amounts to data reduction. This would prevent a fine-grained investigation of differences between mothers' and fathers' unique roles in child-rearing. Moreover, the longitudinal models we used are not suitable for categorical variables, such as parenting style. Still, we have to acknowledge the possibility that the constellation of parenting behaviors is more important than each dimension in isolation. Finally, one potential limitation is that we only included quantitative involvement in the first year of the study. This was done for pragmatic reasons: Children's developmental needs change rapidly, especially in the first years of life. It would be impossible to establish measurement invariance for quantitative involvement from age 6 to 12. Our approach assumes that a "snapshot" of quantitative involvement at age 6 taps into parents' general level of involvement. Consequently, we are unable to assess whether within-person fluctuations in qualitative and quantitative involvement interact. Future research intending to examine such interactions at the within-family level might instead choose to focus on a more restricted age range. A final potential limitation pertains to the generalizability of the findings. From the total sample, we selected our target sample of intact, biological families of natural-born, non-minority, `r masked("Australian")` citizens. It is possible, but beyond the scope of this study, that the findings might have limited generalizability for Australian minorities, or for families in other industrialized nations.

## Conclusions

To recapitulate, we set out to investigate how fathers' and mothers' quantitative and qualitative involvement relate to children's emotional adjustment development. We used innovative methods to disentangle stable between-family differences from within-family fluctuations in qualitative parenting and emotional adjustment. Our results suggest that between-family differences in qualitative parenting behaviors explain a substantial amount of between-family differences in children's emotional adjustment. Moreover, hypotheses based on prior literature that did not disentangle between- and within-family variance were often supported at only one of these levels. This illustrates that analyses aggregating these two levels should not be over-interpreted as reflecting within-family processes.

We found only limited evidence for within-family parenting effects that were significant on average, across families. Even at a relatively young age, child effects outweighed parenting effects in number and size. The direction of effects thus appears to flow primarily from children to parents. We did find that parenting effects varied substantially between families. Some of this variance was explained by between-family differences in joint and relative quantitative involvement: In families where mothers were more involved than fathers, there was greater differentiation between the roles mothers' versus fathers' qualitative parenting played in children's emotional adjustment development. In such families, maternal warmth was more strongly predictive of, and less dependent on, children's emotional adjustment, and fathers' inductive reasoning - a disciplinary strategy - was more contingent on children's emotional adjustment. This suggests that, when differential involvement was higher, parents adopted more differentiated roles: Mothers' role more strongly resembled that of the classic warm "attachment figure", and fathers that of an authority figure. When both parents were more involved, mothers' warmth was less strong of a predictor for emotional adjustment, and mothers' inductive reasoning also became more contingent on adjustment, suggesting that mothers in such families played more similar roles to fathers.

We also found substantial evidence that, when joint involvement was higher, both parents engaged more in parenting behaviors positively associated with children's adjustment. Conversely, when mothers were more involved than fathers, mothers' expressed more warmth, but fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth). Children's emotional adjustment, too, was lower when mothers were relatively more involved than fathers. These findings highlight the important of considering interactions between quantitative and qualitative involvement, and suggest that all family members might benefit from a more equal distribution of childcare tasks.

With regard to mothers' versus fathers' unique roles in children's emotional development, we found that, in line with the notion of the mother as a warm, supportive attachment figure, maternal warmth and consistency were associated with children's emotional adjustment. Evidence for a consistent unique role for fathers was limited, although within-family correlations between paternal inductive reasoning and emotional adjustment suggest that fathers might play a role as disciplinarian. Fathers' parenting effects varied significantly more across families than mothers', however. Combined with the aforementioned interaction effects between qualitative and quantitative involvement, this suggests that - above all - the role of fathers is more context dependent than that of mothers.

\newpage

# References
```{r create_r-references}
#r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
\setlength{\parskip}{0em}

<div id = "refs"></div>
\endgroup