### 2020-05-25 test2.Rmd

---
title: "test"
author: "Caspar J. van Lissa"
date: "3/27/2020"
output: html_document
runtime: shiny
---

### Select car(s)
```{r, echo=FALSE}

## import car data
data <- head(mtcars[1:6])
data$id <- seq(1:nrow(data))

## generate random car prices
set.seed(1)
data$price <- round(runif(nrow(data),5000,15000),0)

myZ <- reactive({
    addCheckboxButtons <- paste0('<input class = "mycheckbox" type="checkbox" name="row', data$id, '" value="op', data$id, '">',"")
  #Display table with checkbox buttons
  z <- cbind(Pick=addCheckboxButtons, data)
  z
})
```

```{r}
output$mytable = renderDataTable({
  myZ()
},
    callback = "function(table) {
    table.on('click.dt', 'tr', function() {
    Shiny.onInputChange('rows',$('.mycheckbox').map(function(){return this.checked*1;}).get().join(\",\"))
    });
}")

## display table
dataTableOutput({"mytable"})
```